---
title: "PenEstim"
subtitle: "Simulation Study"
format: 
  html:
    toc: true
    toc-location: body
date: 4-24-2024
theme: lightly
author:
  - name: "Nicolas Kubista"
    affiliations: "Harvard University"
embed-resources: true
toc: true
toc-expand: 2
toc-title: Contents
page-layout: full
code-fold: true
editor: visual
---

```{r, echo=FALSE, include=FALSE}
library(tidyverse)
library(parallel)
library(dplyr)
library(parallel)
library(stats4)
library(clipp)
library(survival)
library(plyr) #need to load plyr before dplyr
library(truncnorm)
library(PanelPRO)
library(tidyverse)
library(stringr)
library(PedUtils)
library(survival)
library(MASS)
library(profvis)
library(survminer)
library(ggridges)

source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/prior_elicitation.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/helpers.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/outputHelpers.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mhLoglikehood_clipp.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mcmcPenetrance.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mcmcPenetrance_v2.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mcmcPenetrance_v3.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/InternalVars.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Additional Scripts/EstimEval.R")

set.seed(2024)
```

# Family Simulation 1

We simulate 3000 families, our of which we subset the families to only include those families where the proband was a carrier of the PV.
The simulation was done using the PedUtils package. The penetrance for non-carriers is based on the SEER penetrances in the PanelPRO Database. For carriers we specify a Weibull penetrance with the following parameters:

-   Alpha = 2.5

-   Beta = 50

-   Delta = 0

-   Gamma = 1

For men: 

-   Alpha = 2.5

-   Beta = 80

-   Delta = 0

-   Gamma = 1


Censoring is based on a $$N(85,10)$$. The simulated age of death is centered at 85 with a standard deviation of 10, but values above ageMax (default = 94) are re-simulated to be centered around ageMax - 15.

```{r desc fams, echo=TRUE, cache=TRUE}

# Function to describe the families
describeFamilies <- function(fams, cancer, gene) {
  # Validate and translate cancer type
  if (!cancer %in% CANCER_NAME_MAP$long) {
    stop("Cancer type '", cancer, "' is not supported. Please choose from the supported list.")
  }
  
  # Get the abbreviation for the cancer type
  cancer_index <- which(CANCER_NAME_MAP$long == cancer)
  cancer_short <- CANCER_NAME_MAP$short[cancer_index]
  
  # Construct column names based on cancer type and gene input
  aff_col_name <- paste0("isAff", cancer_short)
  age_col_name <- paste0("Age", cancer_short)
  
  # Initialize counters and vectors for storing data
  affectedFamilies = 0 
  affectedProbands = 0
  famSizes = c()
  geneFamilies = 0
  geneProbands = 0 
  curAges = c()
  cancerAges = c()
  pbCancerAges = c()
  pbCurAges = c()
  pbCancerAgesFemale = c()
  affectedFemaleProbands = 0
  geneFemaleProbands = 0 
  cancerAgesFemale = c()
  total_individuals = 0
  gene_individuals = 0
  
  # Process each family
  for(i in 1:length(fams)){
    f = fams[[i]]
    sizeOfFamily = nrow(f)
    total_individuals = total_individuals + sizeOfFamily
    famSizes = c(famSizes, sizeOfFamily)
    curAges = c(curAges, f$CurAge)  
    f = f %>% dplyr::filter(!!rlang::sym(aff_col_name) == 1, isProband == 0) 
    cancerAges <- c(cancerAges, f[[age_col_name]])  
    ff = fams[[i]] %>% dplyr::filter(!!rlang::sym(gene) == 1, !!rlang::sym(aff_col_name) == 1) 
    gene_individuals = gene_individuals + nrow(ff)
    cancerAgesFemale <- c(cancerAgesFemale, ff[[age_col_name]])  

    if(nrow(f)>0){
      affectedFamilies = affectedFamilies + 1
    }
    geneFams = fams[[i]] %>% dplyr::filter(isProband == 0, !!rlang::sym(gene) == 1)
    if(nrow(geneFams) > 0){
     geneFamilies = geneFamilies + 1
    }
    
    pb <- fams[[i]] %>% dplyr::filter(isProband == 1)
    pbfemale <- fams[[i]] %>% dplyr::filter(isProband == 1, Sex == 0)
    
    if (nrow(pbfemale) > 0) {
      pbCancerAgesFemale <- c(pbCancerAgesFemale, pbfemale[[age_col_name]])
      if (any(pbfemale[[aff_col_name]] == 1)){
        affectedFemaleProbands <- affectedFemaleProbands + 1
      }
      if (!is.na(pbfemale[[gene]])){
        if (any(pbfemale[[gene]] == 1)){
          geneFemaleProbands <- geneFemaleProbands + 1
        }
      }
    }
  
    pbCancerAges <- c(pbCancerAges, pb[[age_col_name]])
    pbCurAges <- c(pbCurAges, pb$CurAge)  
    
    if (any(pb[[aff_col_name]] == 1)){
      affectedProbands <- affectedProbands + 1
    }
    
    if (!is.na(pb[[gene]])){
      if (any(pb[[gene]] == 1)){
        geneProbands <- geneProbands + 1
      }
    }
  }

  print(paste0("Number of families: ", length(fams)))
  print(paste0("Total Individuals: ", total_individuals))
  print(paste0("Average family size: ", mean(famSizes)))
  print("Summary of family sizes")
  print(summary(famSizes))
  print("Summary of Current Age for all family members")
  print(summary(curAges))
  print("Summary of Current Age for probands")
  print(summary(pbCurAges))
  
  print(paste0("Number of families with affected individuals (", cancer, "): ", 
               affectedFamilies))
  print(paste0("Number of families with affected probands (", cancer, "): ", 
               affectedProbands))
  print(paste0("Number of families with affected female probands (", cancer, "): ", 
               affectedFemaleProbands))
  print(paste0("Number of families with relatives with PV (and not the proband) (", gene, "): ", geneFamilies))
  print(paste0("Number of probands with PV (", gene, "): ", geneProbands))
  print(paste0("Number of female probands with PV (", gene, "): ", 
               geneFemaleProbands))
  print(paste0("Number of individuals with PV (", gene, "): ", gene_individuals))
  
  print("Summary of Cancer Age of (female) probands")
  print(summary(pbCancerAgesFemale))

  print("Summary of Cancer Age in affected (non-proband) individuals")
  print(summary(cancerAges))
}


```


```{r fam4, echo=FALSE, cache=TRUE}
#set seed
set.seed(2024)

# Generate Families
numberFamilies <- 3000  # Set number of families to be generated
families_sim4 <- list()  # Initialize list for families data
probandIDS = c()
probandBRCA1Status = c()
probandAffectionStatus = c()

# Set up age sequence as in the PanelPRO database
age <- 1:94

# Function to deeply clone a database
deep_clone <- function(x) {
  rapply(x, function(y) if(is.list(y)) return(deep_clone(y)) else return(y), how = "replace")
}

# Clone the original database to create a new simulation database using base R
db_sim <- deep_clone(PanelPRO::PanelPRODatabase)

# Set up allele frequencies
BRCA1freq <- 0.1  # use same BRCA1 frequency (not estimated)
db_sim$AlleleFrequency[["BRCA1_anyPV", 1]] <- BRCA1freq
db_sim$AlleleFrequency[["BRCA1_anyPV", 2]] <- BRCA1freq
db_sim$AlleleFrequency[["BRCA1_anyPV", 3]] <- BRCA1freq

  alpha <- 2.5
  beta <- 50
  gamma <- 1
  delta <- 0
  penetrance.mod <- dweibull(age - delta, alpha, beta) * gamma
  
  dim_names <- attr(db_sim$Penetrance, "dimnames")
  
  cancer_index <- which(dim_names[[1]] == "Breast")
  gene_index <- which(dim_names[[2]] == "BRCA1_hetero_anyPV")
  sex_index <- which(dim_names[[4]] == "Female")
  type_index <- which(dim_names[[6]] == "Net")
  
  # Creating the array with the correct dimensions to fill
  # The dimensions are Cancer, Gene, Race, Sex, Age, Type
  new_array_dims <- c(1, 1, length(dim_names[[3]]), 1, 94, 1)
  new_array <- array(rep(penetrance.mod, each = 8), dim = new_array_dims)
  
  # Correctly setting the penetrance for specified gene, cancer, across all races and types for the specified sex and age
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array

  # Crude
  sex_index <- which(dim_names[[4]] == "Female")
  type_index <- which(dim_names[[6]] == "Crude")
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array
  
  # For Males
  
  alpha <- 2.5
  beta <- 80
  gamma <- 1
  delta <- 0
  penetrance.mod <- dweibull(age - delta, alpha, beta) * gamma
  
  dim_names <- attr(db_sim$Penetrance, "dimnames")
  
  cancer_index <- which(dim_names[[1]] == "Breast")
  gene_index <- which(dim_names[[2]] == "BRCA1_hetero_anyPV")
  sex_index <- which(dim_names[[4]] == "Male")
  type_index <- which(dim_names[[6]] == "Net")
  
  
  # Creating the array with the correct dimensions to fill
  # The dimensions are Cancer, Gene, Race, Sex, Age, Type
  new_array_dims <- c(1, 1, length(dim_names[[3]]), 1, 94, 1)
  new_array <- array(rep(penetrance.mod, each = 8), dim = new_array_dims)
  
  # Correctly setting the penetrance for specified gene, cancer, across all races and types for the specified sex and age
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array
  
  # Crude
  sex_index <- which(dim_names[[4]] == "Male")
  type_index <- which(dim_names[[6]] == "Crude")
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array
  
# Simulation
 suppressMessages(suppressWarnings(capture.output(for(i in 1:numberFamilies){
  # Cancers
  cancers = "Breast"
  # Genes
  genes = "BRCA1"
  #family members
  # Paternal aunts, paternal uncles
  nSibsPatern =floor(rtruncnorm(n=2, mean=0, 1))
  # Maternal aunts, maternal uncles
  nSibsMatern = floor(rtruncnorm(n=2, mean=0, 1))
  # Sisters and brothers
  nSibs = floor(rtruncnorm(n=2, mean=1, 1))
  # We make the assumption that the number of sons and daughters for the
  # proband and all siblings, is the same. Nieces and nephews of the proband
  # are not sampled separately1
  nGrandchild = floor(rtruncnorm(n=2, mean=0, 1))
  nChild = floor(rtruncnorm(n=2, mean=1, 1))
  
  # Simulate family using `PedUtils` code
  fam = sim.runSimFam(nSibsPatern, nSibsMatern, nSibs, nChild,
                      db_sim, genes, cancers,
                      includeGeno = TRUE, includeBiomarkers = FALSE, 
                      censoring = TRUE)
  
  famDF = as.data.frame(fam)
  proband = famDF %>% filter(isProband==1)
  probandIDS = c(probandIDS, proband$ID)
  probandBRCA1Status = c(probandBRCA1Status, proband$BRCA1)
  probandAffectionStatus = c(probandAffectionStatus, proband$isAffBC)
  families_sim4[[i]] = famDF
  
})))

save(families_sim4, file = "families_sim4.RData")
# Filter families with affected probands
carrierProbandfamilies_sim4 <- Filter(function(fam) 
  any(fam$isProband == 1 & fam$BRCA1 == 1), families_sim4)


# hide genotype information for everyone but the proband
simFamiliesGeno <- function(fams) {
  
  simFamilies <- list()
  # Assuming you have the original list of families stored as `original_families`
  # and a vector of proband IDs called `proband_ids`
  for (i in 1:length(fams)) {
    family <- fams[[i]]
    proband <-  family %>% filter(isProband==1)
    family$BRCA1 <- ifelse(family$ID == proband$ID, family$BRCA1, NA)
    simFamilies[[i]] <- family
  }
  return(simFamilies)
  
}

# Subset the family list
simfamilies_sim4 <- simFamiliesGeno(carrierProbandfamilies_sim4)
save(simfamilies_sim4, file = "simfamilies_sim4.Rdata")
print(describeFamilies(simfamilies_sim4,"Breast","BRCA1"))

# Analysis
# Perform survival analysis stratified by gender
data <- do.call(rbind, carrierProbandfamilies_sim4)
fit <- survfit(Surv(AgeBC, isAffBC) ~ BRCA1 + strata(Sex), data = data)
ggsurvplot(fit, risk.table = TRUE, data = data, palette = "Dark2")

# Modify the 'Sex' column to be a factor with names
data$Sex <- factor(data$Sex, levels = c(0, 1), labels = c("Female", "Male"))

# Plot density ridges for affected individuals, stratified by Sex
  ggplot(data = data[data$isAffBC == 1,], aes(x = AgeBC, y = as.factor(BRCA1), fill = as.factor(BRCA1))) +
  geom_density_ridges() +
  facet_wrap(~ Sex, scales = "free_y") +
   ggtitle("Distribution of AgeBC for Affected Individuals")

# Plot density ridges for non-affected individuals, stratified by Sex

ggplot(data = data[data$isAffBC == 0,], aes(x = AgeBC, y = as.factor(BRCA1), fill = as.factor(BRCA1))) +
geom_density_ridges() +
facet_wrap(~ Sex, scales = "free_y") +
ggtitle("Distribution of AgeBC for Unaffected Individuals")

```

# Family Simulation 2
```{r fam4.2, echo=FALSE, cache=TRUE}
#set seed
set.seed(2024)

# Generate Families
numberFamilies <- 3000  # Set number of families to be generated
families_sim4.2 <- list()  # Initialize list for families data
probandIDS = c()
probandBRCA1Status = c()
probandAffectionStatus = c()

# Set up age sequence as in the PanelPRO database
age <- 1:94

# Function to deeply clone a database
deep_clone <- function(x) {
  rapply(x, function(y) if(is.list(y)) return(deep_clone(y)) else return(y), how = "replace")
}

# Clone the original database to create a new simulation database using base R
db_sim <- deep_clone(PanelPRO::PanelPRODatabase)

# Set up allele frequencies
BRCA1freq <- 0.1  # use same BRCA1 frequency (not estimated)
db_sim$AlleleFrequency[["BRCA1_anyPV", 1]] <- BRCA1freq
db_sim$AlleleFrequency[["BRCA1_anyPV", 2]] <- BRCA1freq
db_sim$AlleleFrequency[["BRCA1_anyPV", 3]] <- BRCA1freq

  alpha <- 2.5
  beta <- 50
  gamma <- 0.9
  delta <- 25
  penetrance.mod <- dweibull(age - delta, alpha, beta) * gamma
  
  dim_names <- attr(db_sim$Penetrance, "dimnames")
  
  cancer_index <- which(dim_names[[1]] == "Breast")
  gene_index <- which(dim_names[[2]] == "BRCA1_hetero_anyPV")
  sex_index <- which(dim_names[[4]] == "Female")
  type_index <- which(dim_names[[6]] == "Net")
  
  # Creating the array with the correct dimensions to fill
  # The dimensions are Cancer, Gene, Race, Sex, Age, Type
  new_array_dims <- c(1, 1, length(dim_names[[3]]), 1, 94, 1)
  new_array <- array(rep(penetrance.mod, each = 8), dim = new_array_dims)
  
  # Correctly setting the penetrance for specified gene, cancer, across all races and types for the specified sex and age
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array

  # Crude
  sex_index <- which(dim_names[[4]] == "Female")
  type_index <- which(dim_names[[6]] == "Crude")
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array
  
  # For Males
  
  alpha <- 2
  beta <- 100
  gamma <- 0.9
  delta <- 25
  penetrance.mod <- dweibull(age - delta, alpha, beta) * gamma
  
  dim_names <- attr(db_sim$Penetrance, "dimnames")
  
  cancer_index <- which(dim_names[[1]] == "Breast")
  gene_index <- which(dim_names[[2]] == "BRCA1_hetero_anyPV")
  sex_index <- which(dim_names[[4]] == "Male")
  type_index <- which(dim_names[[6]] == "Net")
  
  
  # Creating the array with the correct dimensions to fill
  # The dimensions are Cancer, Gene, Race, Sex, Age, Type
  new_array_dims <- c(1, 1, length(dim_names[[3]]), 1, 94, 1)
  new_array <- array(rep(penetrance.mod, each = 8), dim = new_array_dims)
  
  # Correctly setting the penetrance for specified gene, cancer, across all races and types for the specified sex and age
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array
  
  # Crude
  sex_index <- which(dim_names[[4]] == "Male")
  type_index <- which(dim_names[[6]] == "Crude")
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array
  
# Simulation
 suppressMessages(suppressWarnings(capture.output(for(i in 1:numberFamilies){
  # Cancers
  cancers = "Breast"
  # Genes
  genes = "BRCA1"
  #family members
  # Paternal aunts, paternal uncles
  nSibsPatern =floor(rtruncnorm(n=2, mean=0, 1))
  # Maternal aunts, maternal uncles
  nSibsMatern = floor(rtruncnorm(n=2, mean=0, 1))
  # Sisters and brothers
  nSibs = floor(rtruncnorm(n=2, mean=1, 1))
  # We make the assumption that the number of sons and daughters for the
  # proband and all siblings, is the same. Nieces and nephews of the proband
  # are not sampled separately1
  nGrandchild = floor(rtruncnorm(n=2, mean=0, 1))
  nChild = floor(rtruncnorm(n=2, mean=1, 1))
  
  # Simulate family using `PedUtils` code
  fam = sim.runSimFam(nSibsPatern, nSibsMatern, nSibs, nChild,
                      db_sim, genes, cancers,
                      includeGeno = TRUE, includeBiomarkers = FALSE, 
                      censoring = TRUE)
  
  famDF = as.data.frame(fam)
  proband = famDF %>% filter(isProband==1)
  probandIDS = c(probandIDS, proband$ID)
  probandBRCA1Status = c(probandBRCA1Status, proband$BRCA1)
  probandAffectionStatus = c(probandAffectionStatus, proband$isAffBC)
  families_sim4.2[[i]] = famDF
  
})))

save(families_sim4.2, file = "families_sim4.2.RData")
# Filter families with affected probands
carrierProbandfamilies_sim4.2 <- Filter(function(fam) 
  any(fam$isProband == 1 & fam$BRCA1 == 1), families_sim4.2)


# hide genotype information for everyone but the proband
simFamiliesGeno <- function(fams) {
  
  simFamilies <- list()
  # Assuming you have the original list of families stored as `original_families`
  # and a vector of proband IDs called `proband_ids`
  for (i in 1:length(fams)) {
    family <- fams[[i]]
    proband <-  family %>% filter(isProband==1)
    family$BRCA1 <- ifelse(family$ID == proband$ID, family$BRCA1, NA)
    simFamilies[[i]] <- family
  }
  return(simFamilies)
  
}

# Subset the family list
simfamilies_sim4.2 <- simFamiliesGeno(carrierProbandfamilies_sim4.2)
save(simfamilies_sim4.2, file = "simfamilies_sim4.2.Rdata")
print(describeFamilies(simfamilies_sim4.2,"Breast","BRCA1"))

# Analysis
# Perform survival analysis stratified by gender
data <- do.call(rbind, carrierProbandfamilies_sim4.2)
fit <- survfit(Surv(AgeBC, isAffBC) ~ BRCA1 + strata(Sex), data = data)
ggsurvplot(fit, risk.table = TRUE, data = data, palette = "Dark2")

# Modify the 'Sex' column to be a factor with names
data$Sex <- factor(data$Sex, levels = c(0, 1), labels = c("Female", "Male"))

# Plot density ridges for affected individuals, stratified by Sex
  ggplot(data = data[data$isAffBC == 1,], aes(x = AgeBC, y = as.factor(BRCA1), fill = as.factor(BRCA1))) +
  geom_density_ridges() +
  facet_wrap(~ Sex, scales = "free_y") +
   ggtitle("Distribution of AgeBC for Affected Individuals")

# Plot density ridges for non-affected individuals, stratified by Sex

ggplot(data = data[data$isAffBC == 0,], aes(x = AgeBC, y = as.factor(BRCA1), fill = as.factor(BRCA1))) +
geom_density_ridges() +
facet_wrap(~ Sex, scales = "free_y") +
ggtitle("Distribution of AgeBC for Unaffected Individuals")

```


# Estimation


The likelihood model in `penet.fn` is based on the Weibull distribution for carriers and, per default, the SEER peentrance data for non-carriers. For an individual $i$, the model considers the following parameters:

-   $\alpha$: Shape parameter of the Weibull distribution
-   $\beta$: Scale parameter of the Weibull distribution
-   $\delta$: Shift parameter for the age in the Weibull function 
-   $\gamma$: Asymptote parameter to adjust the scale of the Weibull function 
-   $\text{age}_i$: Age of the individual $i$
-   $\text{aff}_i$: Affection status (1 if affected, 0 otherwise)

An indiviudal that develops cancer at $\text{age}_i$ contributes a term $f$ to the likelihood, which is just the density of the Weibull in our model. The contribution from an individual whose survival time is censored at is the survival probability $S_i = 1 - F$ where $F$ is corresponding cumulative distribution function of the Weibull. 

For the function \( f(x; \alpha, \beta, \delta, \gamma) \) we define:

$$
f(x; \alpha, \beta, \delta, \gamma) = 
\begin{cases} 
\gamma \left( \frac{\alpha}{\beta} \left( \frac{x - \delta}{\beta} \right)^{\alpha - 1} e^{-\left( \frac{x - \delta}{\beta} \right)^\alpha } \right), & x \geq \delta \\
0, & x < \delta
\end{cases}
$$

And for the cumulative distribution function \( F(x; \alpha, \beta, \delta, \gamma) \):

$$
F(x; \alpha, \beta, \delta, \gamma) =
\begin{cases} 
\gamma \left( 1 - e^{-\left( \frac{x - \delta}{\beta} \right)^\alpha } \right), & x \geq \delta \\
0, & x < \delta
\end{cases}
$$

The survival probability for individual $i$ is calculated as: $$S_i = 1 - pweibull(\text{age}_i - \delta, \alpha, \beta) \times \gamma$$ where pweibull is the CDF of the Weibull.

To simpflify the estimation we fix $\gamma = 1$ and $\delta = 0$.

The density (for affected individuals) is calculated as:

$$ c.pen = dweibull(\text{age}_i - \delta, \alpha, \beta) \times \gamma $$. This gives:

$$
\text{carrierLik}_i = 
\begin{cases} 
\text{S}_i & \text{aff}_i = 0 \\
\text{c.pen} & \text{aff}_i = 1
\end{cases}
$$

The baseline risk $\text{SEER\_baseline}_i$ is derived from SEER data. For more details see seperate documentation. 


## MCMC with Adaptive MH - Families 1

The Adaptive Metropolis-Hastings (MH) algorithm improves the efficiency of the standard Metropolis-Hastings algorithm by adapting the proposal distribution based on the history of the accepted samples. This then adapts the coveriance matrix.

For the proposal distribution at each step I choose a multivariate normal distribution centered at the current state of the Markov chain. The form of the proposal distribution $q(\mathbf{x} \mid \mathbf{x}_{current})$ for moving from the current state $\mathbf{x}_{current}$ to a new state $\mathbf{x}$ is given by:
\[
q(\mathbf{x} \mid \mathbf{x}_{current}) = \mathcal{N}(\mathbf{x}; \mathbf{x}_{current}, \Sigma),
\]
where $\Sigma$ is the covariance matrix.

The covariance matrix $\Sigma$ is adapted based on the covariance of the accepted samples.

The update rule for the covariance matrix is applied periodically to ensure that the algorithm satisfies the required ergodicity conditions. The update rule can be described as follows:
\begin{enumerate}
    \item Collect all accepted proposals up to the $i$-th iteration into a matrix $\mathbf{A}$.
    \item Compute the empirical covariance matrix $\hat{\Sigma}$ of $\mathbf{A}$.
    \item Update the proposal covariance matrix $\Sigma$ used in the proposal distribution:
    \[
    \Sigma \leftarrow \hat{\Sigma} + \epsilon \mathbf{I},
    \]
    where $\epsilon > 0$ is a small constant added to ensure numerical stability and non-singularity of the covariance matrix, and $\mathbf{I}$ is the identity matrix. Here we assume $\epsilon = 0.001$
\end{enumerate}


```{r est3, echo=TRUE, cache=TRUE}
dat <- simfamilies_sim4

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 900, g2 = 100),
    threshold = list(min = 15, max = 40),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
system.time(out_sim_5.3 <- PenEstim_v3(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 1, n_iter_per_chain = 10000, 
    prior_params = prior_params_BRCA1, af = 0.1, burn_in = 0.1, median_max = FALSE
))
save(out_sim_5.3, file = "out_sim_5.3.Rdata")

# print summary stats
out_sim_5.3$summary_stats
plot_traceSingle(out_sim_5.3$results[[1]])
plot_penetrance_sim(out_sim_5.3$combined_chains, prob = 0.95, max_age = 94,"Female",2.5,50,0,1)

# plot the scatterplot with the median and first quartile
plot(out_sim_5.3$combined_chains$median_male_results, out_sim_5.3$combined_chains$first_quartile_male_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
plot(out_sim_5.3$combined_chains$median_female_results, out_sim_5.3$combined_chains$first_quartile_female_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```


## MCMC with Adaptive MH - Families 2

```{r est4, echo=TRUE, cache=TRUE}
dat <- simfamilies_sim4.2

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 900, g2 = 100),
    threshold = list(min = 15, max = 40),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
system.time(out_sim_5.4 <- PenEstim_v3(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 1, n_iter_per_chain = 10000, 
    prior_params = prior_params_BRCA1, af = 0.1, burn_in = 0.1, median_max = FALSE
))
save(out_sim_5.4, file = "out_sim_5.4.Rdata")

# print summary stats
out_sim_5.4$summary_stats
plot_traceSingle(out_sim_5.4$results[[1]])
plot_penetrance_sim(out_sim_5.4$combined_chains, prob = 0.95, max_age = 94,"Female",2.5,50,0,1)

# plot the scatterplot with the median and first quartile
plot(out_sim_5.4$combined_chains$median_male_results, out_sim_5.4$combined_chains$first_quartile_male_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
plot(out_sim_5.4$combined_chains$median_female_results, out_sim_5.4$combined_chains$first_quartile_female_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```

