---
title: "PenEstim"
subtitle: "Simulation Study"
format: 
  html:
    toc: true
    toc-location: body
date: 5-03-2024
theme: lightly
author:
  - name: "Nicolas Kubista"
    affiliations: "Harvard University"
embed-resources: true
toc: true
toc-expand: 2
toc-title: Contents
page-layout: full
code-fold: true
editor: visual
---

```{r, echo=FALSE, include=FALSE}
library(tidyverse)
library(parallel)
library(dplyr)
library(parallel)
library(stats4)
library(clipp)
library(survival)
library(plyr) #need to load plyr before dplyr
library(truncnorm)
library(PanelPRO)
library(tidyverse)
library(stringr)
library(PedUtils)
library(survival)
library(MASS)
library(profvis)
library(survminer)
library(ggridges)

source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/prior_elicitation.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/helpers.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/outputHelpers.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mhLoglikehood_clipp.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mcmcPenetrance.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mcmcPenetrance_v2.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mcmcPenetrance_v3.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mcmcPenetrance_v4.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mcmcPenetrance_v5.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mcmcPenetrance_v5_fixed.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mcmcPenetrance_v7.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/InternalVars.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Additional Scripts/EstimEval.R")

set.seed(2024)
```

# Family Simulation 1

We simulate 3000 families, our of which we subset the families to only include those families where the proband was a carrier of the PV.
The simulation was done using the PedUtils package. The penetrance for non-carriers is based on the SEER penetrances in the PanelPRO Database. For carriers we specify a Weibull penetrance with the following parameters:

-   Alpha = 2.5

-   Beta = 50

-   Delta = 25

-   Gamma = 1

For men: 

-   Alpha = 2.5

-   Beta = 40

-   Delta = 25

-   Gamma = 1


Censoring is based on a $$N(85,10)$$. The simulated age of death is centered at 85 with a standard deviation of 10, but values above ageMax (default = 94) are re-simulated to be centered around ageMax - 15.

```{r desc fams, echo=TRUE, cache=TRUE}

# Function to describe the families
describeFamilies <- function(fams, cancer, gene) {
  # Validate and translate cancer type
  if (!cancer %in% CANCER_NAME_MAP$long) {
    stop("Cancer type '", cancer, "' is not supported. Please choose from the supported list.")
  }
  
  # Get the abbreviation for the cancer type
  cancer_index <- which(CANCER_NAME_MAP$long == cancer)
  cancer_short <- CANCER_NAME_MAP$short[cancer_index]
  
  # Construct column names based on cancer type and gene input
  aff_col_name <- paste0("isAff", cancer_short)
  age_col_name <- paste0("Age", cancer_short)
  
  # Initialize counters and vectors for storing data
  affectedFamilies = 0 
  affectedProbands = 0
  famSizes = c()
  geneFamilies = 0
  geneProbands = 0 
  curAges = c()
  cancerAges = c()
  pbCancerAges = c()
  pbCurAges = c()
  pbCancerAgesFemale = c()
  affectedFemaleProbands = 0
  geneFemaleProbands = 0 
  cancerAgesFemale = c()
  total_individuals = 0
  gene_individuals = 0
  
  # Process each family
  for(i in 1:length(fams)){
    f = fams[[i]]
    sizeOfFamily = nrow(f)
    total_individuals = total_individuals + sizeOfFamily
    famSizes = c(famSizes, sizeOfFamily)
    curAges = c(curAges, f$CurAge)  
    f = f %>% dplyr::filter(!!rlang::sym(aff_col_name) == 1, isProband == 0) 
    cancerAges <- c(cancerAges, f[[age_col_name]])  
    ff = fams[[i]] %>% dplyr::filter(!!rlang::sym(gene) == 1, !!rlang::sym(aff_col_name) == 1) 
    gene_individuals = gene_individuals + nrow(ff)
    cancerAgesFemale <- c(cancerAgesFemale, ff[[age_col_name]])  

    if(nrow(f)>0){
      affectedFamilies = affectedFamilies + 1
    }
    geneFams = fams[[i]] %>% dplyr::filter(isProband == 0, !!rlang::sym(gene) == 1)
    if(nrow(geneFams) > 0){
     geneFamilies = geneFamilies + 1
    }
    
    pb <- fams[[i]] %>% dplyr::filter(isProband == 1)
    pbfemale <- fams[[i]] %>% dplyr::filter(isProband == 1, Sex == 0)
    
    if (nrow(pbfemale) > 0) {
      pbCancerAgesFemale <- c(pbCancerAgesFemale, pbfemale[[age_col_name]])
      if (any(pbfemale[[aff_col_name]] == 1)){
        affectedFemaleProbands <- affectedFemaleProbands + 1
      }
      if (!is.na(pbfemale[[gene]])){
        if (any(pbfemale[[gene]] == 1)){
          geneFemaleProbands <- geneFemaleProbands + 1
        }
      }
    }
  
    pbCancerAges <- c(pbCancerAges, pb[[age_col_name]])
    pbCurAges <- c(pbCurAges, pb$CurAge)  
    
    if (any(pb[[aff_col_name]] == 1)){
      affectedProbands <- affectedProbands + 1
    }
    
    if (!is.na(pb[[gene]])){
      if (any(pb[[gene]] == 1)){
        geneProbands <- geneProbands + 1
      }
    }
  }

  print(paste0("Number of families: ", length(fams)))
  print(paste0("Total Individuals: ", total_individuals))
  print(paste0("Average family size: ", mean(famSizes)))
  print("Summary of family sizes")
  print(summary(famSizes))
  print("Summary of Current Age for all family members")
  print(summary(curAges))
  print("Summary of Current Age for probands")
  print(summary(pbCurAges))
  
  print(paste0("Number of families with affected individuals (", cancer, "): ", 
               affectedFamilies))
  print(paste0("Number of families with affected probands (", cancer, "): ", 
               affectedProbands))
  print(paste0("Number of families with affected female probands (", cancer, "): ", 
               affectedFemaleProbands))
  print(paste0("Number of families with relatives with PV (and not the proband) (", gene, "): ", geneFamilies))
  print(paste0("Number of probands with PV (", gene, "): ", geneProbands))
  print(paste0("Number of female probands with PV (", gene, "): ", 
               geneFemaleProbands))
  print(paste0("Number of individuals with PV (", gene, "): ", gene_individuals))
  
  print("Summary of Cancer Age of (female) probands")
  print(summary(pbCancerAgesFemale))

  print("Summary of Cancer Age in affected (non-proband) individuals")
  print(summary(cancerAges))
}


```


```{r fam4, echo=FALSE, cache=TRUE}
#set seed
set.seed(2024)

# Generate Families
numberFamilies <- 3000  # Set number of families to be generated
families_sim4 <- list()  # Initialize list for families data
probandIDS = c()
probandBRCA1Status = c()
probandAffectionStatus = c()

# Set up age sequence as in the PanelPRO database
age <- 1:94

# Function to deeply clone a database
deep_clone <- function(x) {
  rapply(x, function(y) if(is.list(y)) return(deep_clone(y)) else return(y), how = "replace")
}

# Clone the original database to create a new simulation database using base R
db_sim <- deep_clone(PanelPRO::PanelPRODatabase)

# Set up allele frequencies
BRCA1freq <- 0.1  # use same BRCA1 frequency (not estimated)
db_sim$AlleleFrequency[["BRCA1_anyPV", 1]] <- BRCA1freq
db_sim$AlleleFrequency[["BRCA1_anyPV", 2]] <- BRCA1freq
db_sim$AlleleFrequency[["BRCA1_anyPV", 3]] <- BRCA1freq

  alpha <- 2.5
  beta <- 50
  gamma <- 1
  delta <- 25
  penetrance.mod <- dweibull(age - delta, alpha, beta) * gamma
  
  dim_names <- attr(db_sim$Penetrance, "dimnames")
  
  cancer_index <- which(dim_names[[1]] == "Breast")
  gene_index <- which(dim_names[[2]] == "BRCA1_hetero_anyPV")
  sex_index <- which(dim_names[[4]] == "Female")
  type_index <- which(dim_names[[6]] == "Net")
  
  # Creating the array with the correct dimensions to fill
  # The dimensions are Cancer, Gene, Race, Sex, Age, Type
  new_array_dims <- c(1, 1, length(dim_names[[3]]), 1, 94, 1)
  new_array <- array(rep(penetrance.mod, each = 8), dim = new_array_dims)
  
  # Correctly setting the penetrance for specified gene, cancer, across all races and types for the specified sex and age
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array

  # Crude
  sex_index <- which(dim_names[[4]] == "Female")
  type_index <- which(dim_names[[6]] == "Crude")
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array
  
  # For Males
  
  alpha <- 2.5
  beta <- 40
  gamma <- 1
  delta <- 25
  penetrance.mod <- dweibull(age - delta, alpha, beta) * gamma
  
  dim_names <- attr(db_sim$Penetrance, "dimnames")
  
  cancer_index <- which(dim_names[[1]] == "Breast")
  gene_index <- which(dim_names[[2]] == "BRCA1_hetero_anyPV")
  sex_index <- which(dim_names[[4]] == "Male")
  type_index <- which(dim_names[[6]] == "Net")
  
  
  # Creating the array with the correct dimensions to fill
  # The dimensions are Cancer, Gene, Race, Sex, Age, Type
  new_array_dims <- c(1, 1, length(dim_names[[3]]), 1, 94, 1)
  new_array <- array(rep(penetrance.mod, each = 8), dim = new_array_dims)
  
  # Correctly setting the penetrance for specified gene, cancer, across all races and types for the specified sex and age
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array
  
  # Crude
  sex_index <- which(dim_names[[4]] == "Male")
  type_index <- which(dim_names[[6]] == "Crude")
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array
  
# Simulation
 suppressMessages(suppressWarnings(capture.output(for(i in 1:numberFamilies){
  # Cancers
  cancers = "Breast"
  # Genes
  genes = "BRCA1"
  #family members
  # Paternal aunts, paternal uncles
  nSibsPatern =floor(rtruncnorm(n=2, mean=0, 1))
  # Maternal aunts, maternal uncles
  nSibsMatern = floor(rtruncnorm(n=2, mean=0, 1))
  # Sisters and brothers
  nSibs = floor(rtruncnorm(n=2, mean=1, 1))
  # We make the assumption that the number of sons and daughters for the
  # proband and all siblings, is the same. Nieces and nephews of the proband
  # are not sampled separately1
  nGrandchild = floor(rtruncnorm(n=2, mean=0, 1))
  nChild = floor(rtruncnorm(n=2, mean=1, 1))
  
  # Simulate family using `PedUtils` code
  fam = sim.runSimFam(nSibsPatern, nSibsMatern, nSibs, nChild,
                      db_sim, genes, cancers,
                      includeGeno = TRUE, includeBiomarkers = FALSE, 
                      censoring = TRUE)
  
  famDF = as.data.frame(fam)
  proband = famDF %>% filter(isProband==1)
  probandIDS = c(probandIDS, proband$ID)
  probandBRCA1Status = c(probandBRCA1Status, proband$BRCA1)
  probandAffectionStatus = c(probandAffectionStatus, proband$isAffBC)
  families_sim4[[i]] = famDF
  
})))

save(families_sim4, file = "families_sim4.RData")
# Filter families with affected probands
carrierProbandfamilies_sim4 <- Filter(function(fam) 
  any(fam$isProband == 1 & fam$BRCA1 == 1), families_sim4)


# hide genotype information for everyone but the proband
simFamiliesGeno <- function(fams) {
  
  simFamilies <- list()
  # Assuming you have the original list of families stored as `original_families`
  # and a vector of proband IDs called `proband_ids`
  for (i in 1:length(fams)) {
    family <- fams[[i]]
    proband <-  family %>% filter(isProband==1)
    family$BRCA1 <- ifelse(family$ID == proband$ID, family$BRCA1, NA)
    simFamilies[[i]] <- family
  }
  return(simFamilies)
  
}

# Subset the family list
simfamilies_sim4 <- simFamiliesGeno(carrierProbandfamilies_sim4)
save(simfamilies_sim4, file = "simfamilies_sim4.Rdata")
print(describeFamilies(simfamilies_sim4,"Breast","BRCA1"))

# Analysis
# Perform survival analysis stratified by gender
data <- do.call(rbind, carrierProbandfamilies_sim4)
fit <- survfit(Surv(AgeBC, isAffBC) ~ BRCA1 + strata(Sex), data = data)
ggsurvplot(fit, risk.table = TRUE, data = data, palette = "Dark2")

# Modify the 'Sex' column to be a factor with names
data$Sex <- factor(data$Sex, levels = c(0, 1), labels = c("Female", "Male"))

# Plot density ridges for affected individuals, stratified by Sex
  ggplot(data = data[data$isAffBC == 1,], aes(x = AgeBC, y = as.factor(BRCA1), fill = as.factor(BRCA1))) +
  geom_density_ridges() +
  facet_wrap(~ Sex, scales = "free_y") +
   ggtitle("Distribution of AgeBC for Affected Individuals")

# Plot density ridges for non-affected individuals, stratified by Sex

ggplot(data = data[data$isAffBC == 0,], aes(x = AgeBC, y = as.factor(BRCA1), fill = as.factor(BRCA1))) +
geom_density_ridges() +
facet_wrap(~ Sex, scales = "free_y") +
ggtitle("Distribution of AgeBC for Unaffected Individuals")

```

# Family Simulation 2

For carriers we specify a Weibull penetrance with the following parameters:

-   Alpha = 2.5

-   Beta = 40

-   Delta = 10

-   Gamma = 0.8

For men: 

-   Alpha = 2

-   Beta = 50

-   Delta = 10

-   Gamma = 0.8

```{r fam4.2, echo=FALSE, cache=TRUE}
#set seed
set.seed(2024)

# Generate Families
numberFamilies <- 3000  # Set number of families to be generated
families_sim4.2 <- list()  # Initialize list for families data
probandIDS = c()
probandBRCA1Status = c()
probandAffectionStatus = c()

# Set up age sequence as in the PanelPRO database
age <- 1:94

# Function to deeply clone a database
deep_clone <- function(x) {
  rapply(x, function(y) if(is.list(y)) return(deep_clone(y)) else return(y), how = "replace")
}

# Clone the original database to create a new simulation database using base R
db_sim <- deep_clone(PanelPRO::PanelPRODatabase)

# Set up allele frequencies
BRCA1freq <- 0.1  # use same BRCA1 frequency (not estimated)
db_sim$AlleleFrequency[["BRCA1_anyPV", 1]] <- BRCA1freq
db_sim$AlleleFrequency[["BRCA1_anyPV", 2]] <- BRCA1freq
db_sim$AlleleFrequency[["BRCA1_anyPV", 3]] <- BRCA1freq

  alpha <- 2.5
  beta <- 40
  gamma <- 0.8
  delta <- 10
  penetrance.mod <- dweibull(age - delta, alpha, beta) * gamma
  
  dim_names <- attr(db_sim$Penetrance, "dimnames")
  
  cancer_index <- which(dim_names[[1]] == "Breast")
  gene_index <- which(dim_names[[2]] == "BRCA1_hetero_anyPV")
  sex_index <- which(dim_names[[4]] == "Female")
  type_index <- which(dim_names[[6]] == "Net")
  
  # Creating the array with the correct dimensions to fill
  # The dimensions are Cancer, Gene, Race, Sex, Age, Type
  new_array_dims <- c(1, 1, length(dim_names[[3]]), 1, 94, 1)
  new_array <- array(rep(penetrance.mod, each = 8), dim = new_array_dims)
  
  # Correctly setting the penetrance for specified gene, cancer, across all races and types for the specified sex and age
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array

  # Crude
  sex_index <- which(dim_names[[4]] == "Female")
  type_index <- which(dim_names[[6]] == "Crude")
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array
  
  # For Males
  
  alpha <- 2
  beta <- 50
  gamma <- 0.8
  delta <- 10
  penetrance.mod <- dweibull(age - delta, alpha, beta) * gamma
  
  dim_names <- attr(db_sim$Penetrance, "dimnames")
  
  cancer_index <- which(dim_names[[1]] == "Breast")
  gene_index <- which(dim_names[[2]] == "BRCA1_hetero_anyPV")
  sex_index <- which(dim_names[[4]] == "Male")
  type_index <- which(dim_names[[6]] == "Net")
  
  
  # Creating the array with the correct dimensions to fill
  # The dimensions are Cancer, Gene, Race, Sex, Age, Type
  new_array_dims <- c(1, 1, length(dim_names[[3]]), 1, 94, 1)
  new_array <- array(rep(penetrance.mod, each = 8), dim = new_array_dims)
  
  # Correctly setting the penetrance for specified gene, cancer, across all races and types for the specified sex and age
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array
  
  # Crude
  sex_index <- which(dim_names[[4]] == "Male")
  type_index <- which(dim_names[[6]] == "Crude")
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array
  
# Simulation
 suppressMessages(suppressWarnings(capture.output(for(i in 1:numberFamilies){
  # Cancers
  cancers = "Breast"
  # Genes
  genes = "BRCA1"
  #family members
  # Paternal aunts, paternal uncles
  nSibsPatern =floor(rtruncnorm(n=2, mean=0, 1))
  # Maternal aunts, maternal uncles
  nSibsMatern = floor(rtruncnorm(n=2, mean=0, 1))
  # Sisters and brothers
  nSibs = floor(rtruncnorm(n=2, mean=1, 1))
  # We make the assumption that the number of sons and daughters for the
  # proband and all siblings, is the same. Nieces and nephews of the proband
  # are not sampled separately1
  nGrandchild = floor(rtruncnorm(n=2, mean=0, 1))
  nChild = floor(rtruncnorm(n=2, mean=1, 1))
  
  # Simulate family using `PedUtils` code
  fam = sim.runSimFam(nSibsPatern, nSibsMatern, nSibs, nChild,
                      db_sim, genes, cancers,
                      includeGeno = TRUE, includeBiomarkers = FALSE, 
                      censoring = TRUE)
  
  famDF = as.data.frame(fam)
  proband = famDF %>% filter(isProband==1)
  probandIDS = c(probandIDS, proband$ID)
  probandBRCA1Status = c(probandBRCA1Status, proband$BRCA1)
  probandAffectionStatus = c(probandAffectionStatus, proband$isAffBC)
  families_sim4.2[[i]] = famDF
  
})))

save(families_sim4.2, file = "families_sim4.2.RData")
# Filter families with affected probands
carrierProbandfamilies_sim4.2 <- Filter(function(fam) 
  any(fam$isProband == 1 & fam$BRCA1 == 1), families_sim4.2)


# hide genotype information for everyone but the proband
simFamiliesGeno <- function(fams) {
  
  simFamilies <- list()
  # Assuming you have the original list of families stored as `original_families`
  # and a vector of proband IDs called `proband_ids`
  for (i in 1:length(fams)) {
    family <- fams[[i]]
    proband <-  family %>% filter(isProband==1)
    family$BRCA1 <- ifelse(family$ID == proband$ID, family$BRCA1, NA)
    simFamilies[[i]] <- family
  }
  return(simFamilies)
  
}

# Subset the family list
simfamilies_sim4.2 <- simFamiliesGeno(carrierProbandfamilies_sim4.2)
save(simfamilies_sim4.2, file = "simfamilies_sim4.2.Rdata")
print(describeFamilies(simfamilies_sim4.2,"Breast","BRCA1"))

# Analysis
# Perform survival analysis stratified by gender
data <- do.call(rbind, carrierProbandfamilies_sim4.2)
fit <- survfit(Surv(AgeBC, isAffBC) ~ BRCA1 + strata(Sex), data = data)
ggsurvplot(fit, risk.table = TRUE, data = data, palette = "Dark2")

# Modify the 'Sex' column to be a factor with names
data$Sex <- factor(data$Sex, levels = c(0, 1), labels = c("Female", "Male"))

# Plot density ridges for affected individuals, stratified by Sex
  ggplot(data = data[data$isAffBC == 1,], aes(x = AgeBC, y = as.factor(BRCA1), fill = as.factor(BRCA1))) +
  geom_density_ridges() +
  facet_wrap(~ Sex, scales = "free_y") +
   ggtitle("Distribution of AgeBC for Affected Individuals")

# Plot density ridges for non-affected individuals, stratified by Sex

ggplot(data = data[data$isAffBC == 0,], aes(x = AgeBC, y = as.factor(BRCA1), fill = as.factor(BRCA1))) +
geom_density_ridges() +
facet_wrap(~ Sex, scales = "free_y") +
ggtitle("Distribution of AgeBC for Unaffected Individuals")

```

# Family Simulation 3

For carriers we specify a Weibull penetrance with the following parameters for both sexes:

-   Alpha = 2.5

-   Beta = 60

-   Delta = 10

-   Gamma = 0.5

```{r fam4.3, echo=FALSE, cache=TRUE}
#set seed
set.seed(2024)

# Generate Families
numberFamilies <- 3000  # Set number of families to be generated
families_sim4.3 <- list()  # Initialize list for families data
probandIDS = c()
probandBRCA1Status = c()
probandAffectionStatus = c()

# Set up age sequence as in the PanelPRO database
age <- 1:94

# Function to deeply clone a database
deep_clone <- function(x) {
  rapply(x, function(y) if(is.list(y)) return(deep_clone(y)) else return(y), how = "replace")
}

# Clone the original database to create a new simulation database using base R
db_sim <- deep_clone(PanelPRO::PanelPRODatabase)

# Set up allele frequencies
BRCA1freq <- 0.1  # use same BRCA1 frequency (not estimated)
db_sim$AlleleFrequency[["BRCA1_anyPV", 1]] <- BRCA1freq
db_sim$AlleleFrequency[["BRCA1_anyPV", 2]] <- BRCA1freq
db_sim$AlleleFrequency[["BRCA1_anyPV", 3]] <- BRCA1freq

  alpha <- 2.5
  beta <- 60
  gamma <- 0.5
  delta <- 10
  penetrance.mod <- dweibull(age - delta, alpha, beta) * gamma
  
  dim_names <- attr(db_sim$Penetrance, "dimnames")
  
  cancer_index <- which(dim_names[[1]] == "Breast")
  gene_index <- which(dim_names[[2]] == "BRCA1_hetero_anyPV")
  sex_index <- which(dim_names[[4]] == "Female")
  type_index <- which(dim_names[[6]] == "Net")
  
  # Creating the array with the correct dimensions to fill
  # The dimensions are Cancer, Gene, Race, Sex, Age, Type
  new_array_dims <- c(1, 1, length(dim_names[[3]]), 1, 94, 1)
  new_array <- array(rep(penetrance.mod, each = 8), dim = new_array_dims)
  
  # Correctly setting the penetrance for specified gene, cancer, across all races and types for the specified sex and age
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array

  # Crude
  sex_index <- which(dim_names[[4]] == "Female")
  type_index <- which(dim_names[[6]] == "Crude")
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array
  
  # For Males
  alpha <- 2.5
  beta <- 60
  gamma <- 0.5
  delta <- 10
  penetrance.mod <- dweibull(age - delta, alpha, beta) * gamma
  
  dim_names <- attr(db_sim$Penetrance, "dimnames")
  
  cancer_index <- which(dim_names[[1]] == "Breast")
  gene_index <- which(dim_names[[2]] == "BRCA1_hetero_anyPV")
  sex_index <- which(dim_names[[4]] == "Male")
  type_index <- which(dim_names[[6]] == "Net")
  
  
  # Creating the array with the correct dimensions to fill
  # The dimensions are Cancer, Gene, Race, Sex, Age, Type
  new_array_dims <- c(1, 1, length(dim_names[[3]]), 1, 94, 1)
  new_array <- array(rep(penetrance.mod, each = 8), dim = new_array_dims)
  
  # Correctly setting the penetrance for specified gene, cancer, across all races and types for the specified sex and age
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array
  
  # Crude
  sex_index <- which(dim_names[[4]] == "Male")
  type_index <- which(dim_names[[6]] == "Crude")
  db_sim$Penetrance[cancer_index, gene_index, , sex_index, , type_index] <- new_array
  
# Simulation
 suppressMessages(suppressWarnings(capture.output(for(i in 1:numberFamilies){
  # Cancers
  cancers = "Breast"
  # Genes
  genes = "BRCA1"
  #family members
  # Paternal aunts, paternal uncles
  nSibsPatern =floor(rtruncnorm(n=2, mean=0, 1))
  # Maternal aunts, maternal uncles
  nSibsMatern = floor(rtruncnorm(n=2, mean=0, 1))
  # Sisters and brothers
  nSibs = floor(rtruncnorm(n=2, mean=1, 1))
  # We make the assumption that the number of sons and daughters for the
  # proband and all siblings, is the same. Nieces and nephews of the proband
  # are not sampled separately1
  nGrandchild = floor(rtruncnorm(n=2, mean=0, 1))
  nChild = floor(rtruncnorm(n=2, mean=1, 1))
  
  # Simulate family using `PedUtils` code
  fam = sim.runSimFam(nSibsPatern, nSibsMatern, nSibs, nChild,
                      db_sim, genes, cancers,
                      includeGeno = TRUE, includeBiomarkers = FALSE, 
                      censoring = TRUE)
  
  famDF = as.data.frame(fam)
  proband = famDF %>% filter(isProband==1)
  probandIDS = c(probandIDS, proband$ID)
  probandBRCA1Status = c(probandBRCA1Status, proband$BRCA1)
  probandAffectionStatus = c(probandAffectionStatus, proband$isAffBC)
  families_sim4.3[[i]] = famDF
  
})))

save(families_sim4.3, file = "families_sim4.3.RData")
# Filter families with affected probands
carrierProbandfamilies_sim4.3 <- Filter(function(fam) 
  any(fam$isProband == 1 & fam$BRCA1 == 1), families_sim4.3)


# hide genotype information for everyone but the proband
simFamiliesGeno <- function(fams) {
  
  simFamilies <- list()
  # Assuming you have the original list of families stored as `original_families`
  # and a vector of proband IDs called `proband_ids`
  for (i in 1:length(fams)) {
    family <- fams[[i]]
    proband <-  family %>% filter(isProband==1)
    family$BRCA1 <- ifelse(family$ID == proband$ID, family$BRCA1, NA)
    simFamilies[[i]] <- family
  }
  return(simFamilies)
  
}

# Subset the family list
simfamilies_sim4.3 <- simFamiliesGeno(carrierProbandfamilies_sim4.3)
save(simfamilies_sim4.3, file = "simfamilies_sim4.3.Rdata")
print(describeFamilies(simfamilies_sim4.3,"Breast","BRCA1"))

# Analysis
# Perform survival analysis stratified by gender
data <- do.call(rbind, carrierProbandfamilies_sim4.3)
fit <- survfit(Surv(AgeBC, isAffBC) ~ BRCA1 + strata(Sex), data = data)
ggsurvplot(fit, risk.table = TRUE, data = data, palette = "Dark2")

# Modify the 'Sex' column to be a factor with names
data$Sex <- factor(data$Sex, levels = c(0, 1), labels = c("Female", "Male"))

# Plot density ridges for affected individuals, stratified by Sex
  ggplot(data = data[data$isAffBC == 1,], aes(x = AgeBC, y = as.factor(BRCA1), fill = as.factor(BRCA1))) +
  geom_density_ridges() +
  facet_wrap(~ Sex, scales = "free_y") +
   ggtitle("Distribution of AgeBC for Affected Individuals")

# Plot density ridges for non-affected individuals, stratified by Sex

ggplot(data = data[data$isAffBC == 0,], aes(x = AgeBC, y = as.factor(BRCA1), fill = as.factor(BRCA1))) +
geom_density_ridges() +
facet_wrap(~ Sex, scales = "free_y") +
ggtitle("Distribution of AgeBC for Unaffected Individuals")

```

# Estimation

The likelihood model in `penet.fn` is based on the Weibull distribution for carriers and, per default, the SEER peentrance data for non-carriers. For an individual $i$, the model considers the following parameters:

-   $\alpha$: Shape parameter of the Weibull distribution
-   $\beta$: Scale parameter of the Weibull distribution
-   $\delta$: Shift parameter for the age in the Weibull function 
-   $\gamma$: Asymptote parameter to adjust the scale of the Weibull function 
-   $\text{age}_i$: Age of the individual $i$
-   $\text{aff}_i$: Affection status (1 if affected, 0 otherwise)

An indiviudal that develops cancer at $\text{age}_i$ contributes a term $f$ to the likelihood, which is just the density of the Weibull in our model. The contribution from an individual whose survival time is censored at is the survival probability $S_i = 1 - F$ where $F$ is corresponding cumulative distribution function of the Weibull. 

For the function \( f(x; \alpha, \beta, \delta, \gamma) \) we define:

$$
f(x; \alpha, \beta, \delta, \gamma) = 
\begin{cases} 
\gamma \left( \frac{\alpha}{\beta} \left( \frac{x - \delta}{\beta} \right)^{\alpha - 1} e^{-\left( \frac{x - \delta}{\beta} \right)^\alpha } \right), & x \geq \delta \\
0, & x < \delta
\end{cases}
$$

And for the cumulative distribution function \( F(x; \alpha, \beta, \delta, \gamma) \):

$$
F(x; \alpha, \beta, \delta, \gamma) =
\begin{cases} 
\gamma \left( 1 - e^{-\left( \frac{x - \delta}{\beta} \right)^\alpha } \right), & x \geq \delta \\
0, & x < \delta
\end{cases}
$$

The survival probability for individual $i$ is calculated as: $$S_i = 1 - pweibull(\text{age}_i - \delta, \alpha, \beta) \times \gamma$$ where pweibull is the CDF of the Weibull.

To simpflify the estimation we fix $\gamma = 1$ and $\delta = 0$.

The density (for affected individuals) is calculated as:

$$ c.pen = dweibull(\text{age}_i - \delta, \alpha, \beta) \times \gamma $$. This gives:

$$
\text{carrierLik}_i = 
\begin{cases} 
\text{S}_i & \text{aff}_i = 0 \\
\text{c.pen} & \text{aff}_i = 1
\end{cases}
$$

The baseline risk $\text{SEER\_baseline}_i$ is derived from SEER data. For more details see seperate documentation. 

## Adaptive MH

The Adaptive Metropolis-Hastings (MH) algorithm improves the efficiency of the standard Metropolis-Hastings algorithm by adapting the proposal distribution based on the history of the accepted samples. This adaptation involves updating the covariance matrix according to a specific rule based on the chain's history. The algorithm we implemented is based on Haario et al. (2001).

They choose the proposal density $N(X(t), \Sigma(t))$. Since $X(t)$ is multivariate, we have a multivariate proposal distribution. In the Adaptive Metropolis (AM) sampler, at each iteration of the chain, a candidate value $X^*$ is sampled from a proposal distribution $N(X(t), s_d* \Sigma(t))$, aiming to tune the covariance matrix $\Sigma(t)$ to improve the acceptance rate.

The crucial aspect of the adaptation is how the covariance of the proposal distribution depends on the chain's history. In the AM algorithm, this is addressed by setting $\Sigma(t) \approx s_d \text{cov}(X_0, ..., X_{t-1}) + s_d \epsilon_0 I_d$ after an initial period, where $s_d$ is a parameter that depends only on dimension $d$ and $\epsilon_0$ is a small constant relative to the size of $S$. Here, $I_d$ denotes the $d$-dimensional identity matrix. To begin, we select an arbitrary, strictly positive definite initial covariance $\Sigma_0$, based on our best prior knowledge (which may be quite poor). 
We set $\epsilon_0 = 0.01$ and $s_d = 2.38^2 / d$.
We select an index $t_0 > 0$ for the length of an initial period and define:


$$
\Sigma_t = 
\begin{cases} 
\Sigma_0 & \text{for } t < t_0, \\
s_d \text{cov}(X_0, ..., X_{t-1}) + s_d \epsilon_0 I_d & \text{for } t \geq t_0.
\end{cases}
$$

The choice for the length of the initial segment $t_0 > 0$ is free, but the bigger it is chosen, the more slowly the effect of the adaptation is felt. In a sense, the size of $t_0$ reflects our trust in the initial covariance $\Sigma_0$.
Here we choose $t_0 = \text{burn-in}$ and a for the initial covariance matrix with the variances for the male/female asymptotes (=0), first quartile (=2), median (=5), and threshold (=5):
$$
\Sigma_0=  \begin{bmatrix}
    0.1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    0 & 0.1 & 0 & 0 & 0 & 0 & 0 & 0 \\
    0 & 0 & 2 & 0 & 0 & 0 & 0 & 0 \\
    0 & 0 & 0 & 2 & 0 & 0 & 0 & 0 \\
    0 & 0 & 0 & 0 & 5 & 0 & 0 & 0 \\
    0 & 0 & 0 & 0 & 0 & 5 & 0 & 0 \\
    0 & 0 & 0 & 0 & 0 & 0 & 5 & 0 \\
    0 & 0 & 0 & 0 & 0 & 0 & 0 & 5
\end{bmatrix}
$$

The role of the parameter $\epsilon_0$ is just to ensure that $\Sigma_t$ will not become singular.
As a basic choice for the scaling parameter, we have adopted the value $s_d = \left(\frac{2.4}{\sqrt{d}}\right)^2$ from Gelman et al. (1996), where it was shown that in a certain sense this choice optimizes the mixing properties of the Metropolis search in the case of Gaussian targets and Gaussian proposals

## MCMC with Adaptive MH - Families 1 with 4 Parameter Weibull


```{r est5, echo=TRUE, cache=TRUE}
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Additional Scripts/simFamilies_sim4.RData")

dat <- simfamilies_sim4

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}


# Exploring different priors 
prior_params <- list(
    asymptote = list(g1 = 1, g2 = 1),
    threshold = list(min = 5, max = 30),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
system.time(out_sim_5.4 <- PenEstim_v7(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 1, n_iter_per_chain = 5000, 
    prior_params = prior_params, af = 0.1, burn_in = 0.1, median_max = TRUE
))
save(out_sim_5.4, file = "out_sim_5.4.Rdata")

# print summary stats
out_sim_5.4$summary_stats
plot_traceSingle(out_sim_5.4$results[[1]])
plot_penetrance_sim(out_sim_5.4$combined_chains, prob = 0.95, max_age = 94,"Female",2.5,50,25,1)
plot_penetrance_sim(out_sim_5.4$combined_chains, prob = 0.95, max_age = 94,"Male",2.5,40,25,1)

# plot the scatterplot with the median and first quartile
plot(out_sim_5.4$combined_chains$median_male_results, out_sim_5.4$combined_chains$first_quartile_male_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
plot(out_sim_5.4$combined_chains$median_female_results, out_sim_5.4$combined_chains$first_quartile_female_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```

```{r acf1, echo=TRUE, cache=TRUE}
## ACF Plots 
acf(out_sim_5.4$results[[1]]$first_quartile_male_samples, main = "ACF of First Quartile - Male")
acf(out_sim_5.4$results[[1]]$first_quartile_female_samples, main = "ACF of First Quartile - Female")
acf(out_sim_5.4$results[[1]]$median_male_samples, main = "ACF of Median - Male")
acf(out_sim_5.4$results[[1]]$median_female_samples, main = "ACF of Median - Female")
acf(out_sim_5.4$results[[1]]$threshold_male_samples, main = "ACF of Threshold - Male")
acf(out_sim_5.4$results[[1]]$threshold_female_samples, main = "ACF of Threshold - Female")
```

## MCMC with Adaptive MH - Families 2 with 4 Parameter Weibull


```{r est6, echo=TRUE, cache=TRUE}
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Additional Scripts/simFamilies_sim4.2.RData")

dat <- simfamilies_sim4.2

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}


# Exploring different priors 
prior_params <- list(
    asymptote = list(g1 = 1, g2 = 1),
    threshold = list(min = 5, max = 30),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
system.time(out_sim_5.6 <- PenEstim_v7(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 1, n_iter_per_chain = 7000, 
    prior_params = prior_params, af = 0.1, burn_in = 0.1, median_max = TRUE
))
save(out_sim_5.6, file = "out_sim_5.6.Rdata")

# print summary stats
out_sim_5.6$summary_stats
plot_traceSingle(out_sim_5.6$results[[1]])
plot_penetrance_sim(out_sim_5.6$combined_chains, prob = 0.95, max_age = 94,"Female",2.5,50,25,1)
plot_penetrance_sim(out_sim_5.6$combined_chains, prob = 0.95, max_age = 94,"Male",2.5,40,25,1)

# plot the scatterplot with the median and first quartile
plot(out_sim_5.6$combined_chains$median_male_results, out_sim_5.6$combined_chains$first_quartile_male_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
plot(out_sim_5.6$combined_chains$median_female_results, out_sim_5.6$combined_chains$first_quartile_female_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
plot(out_sim_5.6$combined_chains$median_male_results, out_sim_5.6$combined_chains$asymptote_male_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
plot(out_sim_5.6$combined_chains$median_female_results, out_sim_5.6$combined_chains$asymptote_female_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```

```{r acf2, echo=TRUE, cache=TRUE}
## ACF Plots 
acf(out_sim_5.6$results[[1]]$first_quartile_male_samples, main = "ACF of First Quartile - Male")
acf(out_sim_5.6$results[[1]]$first_quartile_female_samples, main = "ACF of First Quartile - Female")
acf(out_sim_5.6$results[[1]]$median_male_samples, main = "ACF of Median - Male")
acf(out_sim_5.6$results[[1]]$median_female_samples, main = "ACF of Median - Female")
acf(out_sim_5.6$results[[1]]$threshold_male_samples, main = "ACF of Threshold - Male")
acf(out_sim_5.6$results[[1]]$threshold_female_samples, main = "ACF of Threshold - Female")
```



## MCMC with Adaptive MH - Families 3 with 4 Parameter Weibull

Family 3 was generated with a Weibull penetrance that had an asymptote set at $\gamma = 0.5$. 

```{r est8, echo=TRUE, cache=TRUE}

load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Additional Scripts/simFamilies_sim4.3.RData")
dat <- simfamilies_sim4.3

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params <- list(
    asymptote = list(g1 = 1, g2 = 1),
    threshold = list(min = 5, max = 30),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)
# Run Estimation procedure with default prior setting 
# Main Estimation for Female
system.time(out_sim_5.7 <- PenEstim_v7(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 1, n_iter_per_chain = 500, 
    prior_params = prior_params, af = 0.1, burn_in = 0.1, median_max= TRUE, priors = prior_params))

#save(out_sim_5.7, file = "out_sim_5.7.Rdata")

# print summary stats
out_sim_5.7$summary_stats
plot_traceSingle(out_sim_5.7$results[[1]])
plot_penetrance_sim(out_sim_5.7$combined_chains, prob = 0.95, max_age = 94,"Female",2.5,60,10,0.5)
plot_penetrance_sim(out_sim_5.7$combined_chains, prob = 0.95, max_age = 94,"Male",2.5,60,10,0.5)

# plot the scatterplot with the median and first quartile
plot(out_sim_5.7$combined_chains$median_male_results, out_sim_5.7$combined_chains$first_quartile_male_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
plot(out_sim_5.7$combined_chains$median_female_results, out_sim_5.7$combined_chains$first_quartile_female_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
plot(out_sim_5.7$combined_chains$median_male_results, out_sim_5.7$combined_chains$asymptote_male_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
plot(out_sim_5.7$combined_chains$median_female_results, out_sim_5.7$combined_chains$asymptote_female_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin

# Adjust the plot for male data
plot(out_sim_5.7$combined_chains$median_male_results, 
     100 * out_sim_5.7$combined_chains$asymptote_male_results, 
     xlim = c(1, 100), 
     ylim = c(1, 100),
     xlab = "Median Age", 
     ylab = "Asymptote (Scaled)")
abline(a=0, b=1, col="red")  # This draws a red line with slope 1 passing through the origin

# Adjust the plot for female data
plot(out_sim_5.7$combined_chains$median_female_results, 
     100 * out_sim_5.7$combined_chains$asymptote_female_results, 
     xlim = c(58, 70), 
     ylim = c(20, 70),
     xlab = "Median Age", 
     ylab = "Asymptote (Scaled)")
abline(a=0, b=1, col="red")  # This draws a red line with slope 1 passing through the origin

cor(out_sim_5.7$combined_chains$first_quartile_female_results, 
     100 * out_sim_5.7$combined_chains$asymptote_female_results)

```

```{r ac5, echo=TRUE, cache=TRUE}
## ACF Plots 
acf(out_sim_5.7$results[[1]]$first_quartile_male_samples, main = "ACF of First Quartile - Male")
acf(out_sim_5.7$results[[1]]$first_quartile_female_samples, main = "ACF of First Quartile - Female")
acf(out_sim_5.7$results[[1]]$median_male_samples, main = "ACF of Median - Male")
acf(out_sim_5.7$results[[1]]$median_female_samples, main = "ACF of Median - Female")
acf(out_sim_5.7$results[[1]]$threshold_male_samples, main = "ACF of Threshold - Male")
acf(out_sim_5.7$results[[1]]$threshold_female_samples, main = "ACF of Threshold - Female")
```



## MCMC - MH within Gibbs

Family 3 was generated with a Weibull penetrance that had an asymptote set at $\gamma = 0.5$. 

```{r est10, echo=TRUE, cache=TRUE}

load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Additional Scripts/simFamilies_sim4.3.RData")
dat <- simfamilies_sim4.3

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params <- list(
    asymptote = list(g1 = 1, g2 = 1),
    threshold = list(min = 5, max = 30),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)
# Run Estimation procedure with default prior setting 
# Main Estimation for Female
system.time(out_sim_6.0 <- PenEstim_vMWG(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 1, n_iter_per_chain = 10000, 
    prior_params = prior_params, af = 0.1, burn_in = 0.1, median_max= TRUE, priors = prior_params))

#save(out_sim_6.0, file = "out_sim_6.0.Rdata")

# print summary stats
out_sim_6.0$summary_stats
plot_traceSingle(out_sim_6.0$results[[1]])
plot_penetrance_sim(out_sim_6.0$combined_chains, prob = 0.95, max_age = 94,"Female",2.5,60,10,0.5)
plot_penetrance_sim(out_sim_6.0$combined_chains, prob = 0.95, max_age = 94,"Male",2.5,60,10,0.5)

# plot the scatterplot with the median and first quartile
plot(out_sim_6.0$combined_chains$median_male_results, out_sim_6.0$combined_chains$first_quartile_male_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
plot(out_sim_6.0$combined_chains$median_female_results, out_sim_6.0$combined_chains$first_quartile_female_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
plot(out_sim_6.0$combined_chains$median_male_results, out_sim_6.0$combined_chains$asymptote_male_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
plot(out_sim_6.0$combined_chains$median_female_results, out_sim_6.0$combined_chains$asymptote_female_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin

# Adjust the plot for male data
plot(out_sim_6.0$combined_chains$median_male_results, 
     100 * out_sim_6.0$combined_chains$asymptote_male_results, 
     xlim = c(1, 100), 
     ylim = c(1, 100),
     xlab = "Median Age", 
     ylab = "Asymptote (Scaled)")
abline(a=0, b=1, col="red")  # This draws a red line with slope 1 passing through the origin

# Adjust the plot for female data
plot(out_sim_6.0$combined_chains$median_female_results, 
     100 * out_sim_6.0$combined_chains$asymptote_female_results, 
     xlim = c(58, 70), 
     ylim = c(20, 70),
     xlab = "Median Age", 
     ylab = "Asymptote (Scaled)")
abline(a=0, b=1, col="red")  # This draws a red line with slope 1 passing through the origin

cor(out_sim_6.0$combined_chains$first_quartile_female_results, 
     100 * out_sim_6.0$combined_chains$asymptote_female_results)

```

```{r ac2, echo=TRUE, cache=TRUE}
## ACF Plots 
acf(out_sim_6.0$results[[1]]$first_quartile_male_samples, main = "ACF of First Quartile - Male")
acf(out_sim_6.0$results[[1]]$first_quartile_female_samples, main = "ACF of First Quartile - Female")
acf(out_sim_6.0$results[[1]]$median_male_samples, main = "ACF of Median - Male")
acf(out_sim_6.0$results[[1]]$median_female_samples, main = "ACF of Median - Female")
acf(out_sim_6.0$results[[1]]$threshold_male_samples, main = "ACF of Threshold - Male")
acf(out_sim_6.0$results[[1]]$threshold_female_samples, main = "ACF of Threshold - Female")
```

