---
title: "PenEstim Application_v2"
format:
  html:
    theme: cosmo
    toc: true
    number-sections: true
    left:
        - text: "Home"
          href: index.html
        - text: "Section 1"
          href: #section-1
        - text: "Section 2"
          href: #section-2
---

```{r, echo=FALSE, include=FALSE}
library(tidyverse)
library(PPP)
library(parallel)
library(dplyr)
library(stats4)
library(clipp)
library(survival)


source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/prior_elicitation.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/helpers.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/outputHelpers.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mhLoglikehood_clipp.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mcmcPenetrance.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/InternalVars.R")

set.seed(2024)
```

```{r, echo=FALSE, include=FALSE}
describeFamilies <- function(fams, cancer, gene) {
  # Validate and translate cancer type
  if (!cancer %in% CANCER_NAME_MAP$long) {
    stop("Cancer type '", cancer, "' is not supported. Please choose from the supported list.")
  }
  
  # Get the abbreviation for the cancer type
  cancer_index <- which(CANCER_NAME_MAP$long == cancer)
  cancer_short <- CANCER_NAME_MAP$short[cancer_index]
  
  # Construct column names based on cancer type and gene input
  aff_col_name <- paste0("isAff", cancer_short)
  age_col_name <- paste0("Age", cancer_short)
  
  # Initialize counters and vectors for storing data
  affectedFamilies = 0 
  affectedProbands = 0
  famSizes = c()
  geneFamilies = 0
  geneProbands = 0 
  curAges = c()
  cancerAges = c()
  pbCancerAges = c()
  pbCurAges = c()
  pbCancerAgesFemale = c()
  affectedFemaleProbands = 0
  geneFemaleProbands = 0 
  cancerAgesFemale = c()
  total_individuals = 0
  gene_individuals = 0
  
  # Process each family
  for(i in 1:length(fams)){
    f = fams[[i]]
    sizeOfFamily = nrow(f)
    total_individuals = total_individuals + sizeOfFamily
    famSizes = c(famSizes, sizeOfFamily)
    curAges = c(curAges, f$CurAge)  
    f = f %>% dplyr::filter(!!rlang::sym(aff_col_name) == 1, isProband == 0) 
    cancerAges <- c(cancerAges, f[[age_col_name]])  
    ff = fams[[i]] %>% dplyr::filter(!!rlang::sym(gene) == 1, !!rlang::sym(aff_col_name) == 1) 
    gene_individuals = gene_individuals + nrow(ff)
    cancerAgesFemale <- c(cancerAgesFemale, ff[[age_col_name]])  

    if(nrow(f)>0){
      affectedFamilies = affectedFamilies + 1
    }
    geneFams = fams[[i]] %>% dplyr::filter(isProband == 0, !!rlang::sym(gene) == 1)
    if(nrow(geneFams) > 0){
     geneFamilies = geneFamilies + 1
    }
    
    pb <- fams[[i]] %>% dplyr::filter(isProband == 1)
    pbfemale <- fams[[i]] %>% dplyr::filter(isProband == 1, Sex == 0)
    
    if (nrow(pbfemale) > 0) {
      pbCancerAgesFemale <- c(pbCancerAgesFemale, pbfemale[[age_col_name]])
      if (any(pbfemale[[aff_col_name]] == 1)){
        affectedFemaleProbands <- affectedFemaleProbands + 1
      }
      if (!is.na(pbfemale[[gene]])){
        if (any(pbfemale[[gene]] == 1)){
          geneFemaleProbands <- geneFemaleProbands + 1
        }
      }
    }
  
    pbCancerAges <- c(pbCancerAges, pb[[age_col_name]])
    pbCurAges <- c(pbCurAges, pb$CurAge)  
    
    if (any(pb[[aff_col_name]] == 1)){
      affectedProbands <- affectedProbands + 1
    }
    
    if (!is.na(pb[[gene]])){
      if (any(pb[[gene]] == 1)){
        geneProbands <- geneProbands + 1
      }
    }
  }

  print(paste0("Number of families: ", length(fams)))
  print(paste0("Total Individuals: ", total_individuals))
  print(paste0("Average family size: ", mean(famSizes)))
  print("Summary of family sizes")
  print(summary(famSizes))
  print("Summary of Current Age for all family members")
  print(summary(curAges))
  print("Summary of Current Age for probands")
  print(summary(pbCurAges))
  
  print(paste0("Number of families with affected individuals (", cancer, "): ", 
               affectedFamilies))
  print(paste0("Number of families with affected probands (", cancer, "): ", 
               affectedProbands))
  print(paste0("Number of families with affected female probands (", cancer, "): ", 
               affectedFemaleProbands))
  print(paste0("Number of families with relatives with PV (and not the proband) (", gene, "): ", geneFamilies))
  print(paste0("Number of probands with PV (", gene, "): ", geneProbands))
  print(paste0("Number of female probands with PV (", gene, "): ", 
               geneFemaleProbands))
  print(paste0("Number of individuals with PV (", gene, "): ", gene_individuals))
  
  print("Summary of Cancer Age of (female) probands")
  print(summary(pbCancerAgesFemale))

  print("Summary of Cancer Age in affected (non-proband) individuals")
  print(summary(cancerAges))
  
  print("Summary of Cancer Age in affected (non-proband) female individuals")
  print(summary(cancerAgesFemale))
}

```

# Additional Simulations

## Simulation - Default 

```{r, echo=TRUE, cache=TRUE}

# Load the additonal families
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/peelingpearingestim/simFamilies_C_1000_nocen_selected300.RData")

dat<- simFamilies_C_1000_nocen_selected300

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 1, g2 = 1),
    threshold = list(min = 15, max = 25),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
out_sim_10k_simC300 <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 10000, 
    prior_params = prior_params_BRCA1, af = 0.9, burn_in = 0.1, sex = "Female"
)

# print summary stats
out_sim_10k_simC300$summary_stats
plot_traceSingle(out_sim_10k_simC300$results[[1]])
ages <- 1:94
plot_penetrance(out_sim_10k_simC300$combined_chains, prob = 0.95, max_age = 94)
lines(ages, pweibull(ages - 20, 2.5, 50)*0.9, type = "l", lwd=2)

# Plot the running mean 

runFQ<- running_mean(out_sim_10k_simC300$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_10k_simC300$results[[1]]$median_samples)
runThreshold <- running_mean(out_sim_10k_simC300$results[[1]]$threshold_samples)
runAsy <- running_mean(out_sim_10k_simC300$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Threshold Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_10k_simC300$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_10k_simC300$results[[1]]$median_samples)
runVThreshold <- running_variance(out_sim_10k_simC300$results[[1]]$threshold_samples)
runVAsy <- running_variance(out_sim_10k_simC300$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Threshold Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")


  # Autocorrelation plots
acf(out_sim_10k_simC300$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_10k_simC300$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_10k_simC300$results[[1]]$threshold_samples, main = "ACF of Threshold")
acf(out_sim_10k_simC300$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

# plot the scatterplot with the median and first quartile
plot(out_sim_10k_simC300$combined_chains$median_results, out_sim_10k_simC300$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```

## Simulation - Default but fixed median

```{r, echo=TRUE, cache=TRUE}

# Load the additonal families
dat<- simFamilies_C_1000_nocen_selected300

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 900, g2 = 100),
    threshold = list(min = 15, max = 25),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
out_sim_10k_simC300 <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 10000, 
    prior_params = prior_params_BRCA1, af = 0.9, burn_in = 0.1, sex = "Female"
)

# print summary stats
out_sim_10k_simC300$summary_stats
plot_traceSingle(out_sim_10k_simC300$results[[1]])
plot_penetrance(out_sim_10k_simC300$combined_chains, prob = 0.95, max_age = 94)
lines(ages, pweibull(ages - 20, 2.5, 50)*0.9, type = "l", lwd=2)

# Plot the running mean 

runFQ<- running_mean(out_sim_10k_simC300$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_10k_simC300$results[[1]]$median_samples)
runThreshold <- running_mean(out_sim_10k_simC300$results[[1]]$threshold_samples)
runAsy <- running_mean(out_sim_10k_simC300$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of threshold Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_10k_simC300$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_10k_simC300$results[[1]]$median_samples)
runVThreshold <- running_variance(out_sim_10k_simC300$results[[1]]$threshold_samples)
runVAsy <- running_variance(out_sim_10k_simC300$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Threshold Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")


  # Autocorrelation plots
acf(out_sim_10k_simC300$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_10k_simC300$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_10k_simC300$results[[1]]$threshold_samples, main = "ACF of Threshold")
acf(out_sim_10k_simC300$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

# plot the scatterplot with the median and first quartile
plot(out_sim_10k_simC300$combined_chains$median_results, out_sim_10k_simC300$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```

## Simulation - Uncensored Data Default

```{r, echo=TRUE, cache=TRUE}

# Load the additonal families
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/peelingpearingestim/simFamilies_C_1000_selected300.RData")

dat<- simFamilies_C_1000_selected300

print(describeFamilies(dat,"Breast","BRCA1"))

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 1, g2 = 1),
    threshold = list(min = 15, max = 25),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
out_sim_10k_simC300 <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 5000, 
    prior_params = prior_params_BRCA1, af = 0.9, burn_in = 0.1, sex = "Female"
)

# print summary stats
out_sim_10k_simC300$summary_stats
plot_traceSingle(out_sim_10k_simC300$results[[1]])
ages <- 1:94
plot_penetrance(out_sim_10k_simC300$combined_chains, prob = 0.95, max_age = 94)
lines(ages, pweibull(ages - 20, 2.5, 50)*0.9, type = "l", lwd=2)

# Plot the running mean 

runFQ<- running_mean(out_sim_10k_simC300$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_10k_simC300$results[[1]]$median_samples)
runThreshold <- running_mean(out_sim_10k_simC300$results[[1]]$threshold_samples)
runAsy <- running_mean(out_sim_10k_simC300$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Threshold Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_10k_simC300$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_10k_simC300$results[[1]]$median_samples)
runVThreshold <- running_variance(out_sim_10k_simC300$results[[1]]$threshold_samples)
runVAsy <- running_variance(out_sim_10k_simC300$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Threshold Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")


  # Autocorrelation plots
acf(out_sim_10k_simC300$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_10k_simC300$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_10k_simC300$results[[1]]$threshold_samples, main = "ACF of Threshold")
acf(out_sim_10k_simC300$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

# plot the scatterplot with the median and first quartile
plot(out_sim_10k_simC300$combined_chains$median_results, out_sim_10k_simC300$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```

## Simulation - Uncensored Data and fixed asymptote

```{r, echo=TRUE, cache=TRUE}

# Load the additonal families
dat<- simFamilies_C_1000_nocen_selected300

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 900, g2 = 100),
    threshold = list(min = 15, max = 25),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
out_sim_10k_simC300 <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 5000, 
    prior_params = prior_params_BRCA1, af = 0.9, burn_in = 0.1, sex = "Female"
)

# print summary stats
out_sim_10k_simC300$summary_stats
plot_traceSingle(out_sim_10k_simC300$results[[1]])
plot_penetrance(out_sim_10k_simC300$combined_chains, prob = 0.95, max_age = 94)
lines(ages, pweibull(ages - 20, 2.5, 50)*0.9, type = "l", lwd=2)

# Plot the running mean 

runFQ<- running_mean(out_sim_10k_simC300$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_10k_simC300$results[[1]]$median_samples)
runThreshold <- running_mean(out_sim_10k_simC300$results[[1]]$threshold_samples)
runAsy <- running_mean(out_sim_10k_simC300$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of threshold Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_10k_simC300$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_10k_simC300$results[[1]]$median_samples)
runVThreshold <- running_variance(out_sim_10k_simC300$results[[1]]$threshold_samples)
runVAsy <- running_variance(out_sim_10k_simC300$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Threshold Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")


  # Autocorrelation plots
acf(out_sim_10k_simC300$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_10k_simC300$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_10k_simC300$results[[1]]$threshold_samples, main = "ACF of Threshold")
acf(out_sim_10k_simC300$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

# plot the scatterplot with the median and first quartile
plot(out_sim_10k_simC300$combined_chains$median_results, out_sim_10k_simC300$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```

# Application - CoH Data (Hispanic Cohort)

## MLH6 and Colorectal - Default Prior

Used the default prior settings but used a wider threshold parameter due to generally later age of onset.

```{r, echo=TRUE, cache=TRUE}

load("/Users/nicolaskubista/Dropbox (Partners HealthCare)/CCGCRN Hispanic Cohort Data/PenEstim/Data/carrierProbandFamilies_cohPedigree_MSH6_COL.RData")

# Exploring different priors 
prior_params_MSH6 <- list(
    asymptote = list(g1 = 1, g2 = 1),
    threshold = list(min = 0, max = 50), 
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
out_sim_10k_MSH6 <- PenEstim(
    data = carrierProbandFamilies_cohPedigree_MSH6_COL,
    cancer_type = "Colorectal", gene_input = "MSH6", n_chains = 4, n_iter_per_chain = 10000,
    prior_params = prior_params_MSH6,
    af = PanelPRODatabase$AlleleFrequency[["MSH6_anyPV",1]], burn_in = 0.1
)

# print summary stats
out_sim_10k_MSH6$summary_stats
plot_traceSingle(out_sim_10k_MSH6$results[[1]])

# Plot the running mean 

runFQ<- running_mean(out_sim_10k_MSH6$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_10k_MSH6$results[[1]]$median_samples)
runThreshold <- running_mean(out_sim_10k_MSH6$results[[1]]$threshold_samples)
runAsy <- running_mean(out_sim_10k_MSH6$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Threshold Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_10k_MSH6$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_10k_MSH6$results[[1]]$median_samples)
runVThreshold <- running_variance(out_sim_10k_MSH6$results[[1]]$threshold_samples)
runVAsy <- running_variance(out_sim_10k_MSH6$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Threshold Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")

  # Autocorrelation plots
acf(out_sim_10k_MSH6$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_10k_MSH6$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_10k_MSH6$results[[1]]$threshold_samples, main = "ACF of Threshold")
acf(out_sim_10k_MSH6$results[[1]]$asymptote_samples, main = "ACF of Asymptote")
```

## MSH6 and Colorectal - New Prior

Here we added are more skewed prior on the median

```{r, echo=TRUE, cache=TRUE}

# Exploring different priors 
prior_params_MSH6 <- list(
    asymptote = list(g1 = 1, g2 = 1),
    threshold = list(min = 0, max = 50), # new
    median = list(m1 = 8, m2 = 2), # this is new 
    first_quartile = list(q1 = 6, q2 = 3)
)

# New
out_sim_10k_MSH6_newPriorMP <- PenEstim(
    data = carrierProbandFamilies_cohPedigree_MSH6_COL,
    cancer_type = "Colorectal", gene_input = "MSH6", n_chains = 4, n_iter_per_chain = 10000, median_max = FALSE,
    prior_params = prior_params_MSH6, max_penetrance =1,
    af = PanelPRODatabase$AlleleFrequency[["MSH6_anyPV",1]], burn_in = 0.1
)

# print summary stats
out_sim_10k_MSH6_newPriorMP$summary_stats
plot_traceSingle(out_sim_10k_MSH6_newPriorMP$results[[1]])
generate_density_plots(out_sim_10k_MSH6_newPriorMP$combined_chains)

# Plot the running mean 

runFQ<- running_mean(out_sim_10k_MSH6_newPriorMP$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_10k_MSH6_newPriorMP$results[[1]]$median_samples)
runThreshold <- running_mean(out_sim_10k_MSH6_newPriorMP$results[[1]]$threshold_samples)
runAsy <- running_mean(out_sim_10k_MSH6_newPriorMP$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Threshold Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_10k_MSH6_newPriorMP$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_10k_MSH6_newPriorMP$results[[1]]$median_samples)
runVThreshold <- running_variance(out_sim_10k_MSH6_newPriorMP$results[[1]]$threshold_samples)
runVAsy <- running_variance(out_sim_10k_MSH6_newPriorMP$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Threshold Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")

  # Autocorrelation plots
acf(out_sim_10k_MSH6_newPriorMP$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_10k_MSH6_newPriorMP$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_10k_MSH6_newPriorMP$results[[1]]$threshold_samples, main = "ACF of threshold")
acf(out_sim_10k_MSH6_newPriorMP$results[[1]]$asymptote_samples, main = "ACF of Asymptote")


# plot the scatterplot with the median and first quartile
plot(out_sim_10k_MSH6_newPriorMP$combined_chains$median_results, out_sim_10k_MSH6_newPriorMP$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin

```

# Penetrance Estimates from PanelPRO

As a reference, I added the penetrance curves for the respective gene-cancer combinations from the PanelPRODatabase here.


```{r, echo=TRUE, cache=TRUE}
db <- PPP::PanelPRODatabase
ages <- 1:94
# Extracting penetrance data for females
pen1_f <- cumsum(db$Penetrance["Breast", "BRCA1_hetero_anyPV", "All_Races", "Female", ,"Net"])

# Assuming similar structure, extracting for males (modify according to your data structure)
pen1_m <- cumsum(db$Penetrance["Breast", "BRCA1_hetero_anyPV", "All_Races", "Male", ,"Net"])

# Plot for BRCA1 and Breast Cancer including both genders
plot(ages, pen1_f, type="l", lwd=2, col="blue",
     xlab="Age", ylab="Cumulative Penetrance", xlim = c(1,100), ylim = c(0,1),
     main="BRCA1 and Breast Cancer")
lines(ages, pen1_m, type="l", lwd=2, col="red")
legend("topleft", legend=c("Female", "Male"), col=c("blue", "red"), lty=1, lwd=2)

# MLH1 and Colorectal Cancer for both genders
pen2_f <- cumsum(db$Penetrance["Colorectal", "MLH1_hetero_anyPV", "All_Races", "Female", ,"Net"])
pen2_m <- cumsum(db$Penetrance["Colorectal", "MLH1_hetero_anyPV", "All_Races", "Male", ,"Net"])

# Plot for MLH1 and Colorectal Cancer
plot(ages, pen2_f, type="l", lwd=2, col="blue",
     xlab="Age", ylab="Cumulative Penetrance", xlim = c(1,100),ylim = c(0,1),
     main="MLH1 and Colorectal Cancer")
lines(ages, pen2_m, type="l", lwd=2, col="red")
legend("topleft", legend=c("Female", "Male"), col=c("blue", "red"), lty=1, lwd=2)

# MSH6 and Colorectal Cancer for both genders
pen3_f <- cumsum(db$Penetrance["Colorectal", "MSH6_hetero_anyPV", "All_Races", "Female", ,"Net"])
pen3_m <- cumsum(db$Penetrance["Colorectal", "MSH6_hetero_anyPV", "All_Races", "Male", ,"Net"])

# Plot for MSH6 and Colorectal Cancer
plot(ages, pen3_f, type="l", lwd=2, col="blue",
     xlab="Age", ylab="Cumulative Penetrance", xlim = c(1,100),ylim = c(0,1),
     main="MSH6 and Colorectal Cancer")
lines(ages, pen3_m, type="l", lwd=2, col="red")
legend("topleft", legend=c("Female", "Male"), col=c("blue", "red"), lty=1, lwd=2)

```

