---
title: "PenEstim"
subtitle: "Simulation Study"
format: 
  html:
    toc: true
    toc-location: body
date: 3-19-2024
theme: lightly
author:
  - name: "Nicolas Kubista"
    affiliations: "Harvard University"
embed-resources: true
toc: true
toc-expand: 2
toc-title: Contents
page-layout: full
code-fold: true
editor: visual
---

```{r, echo=FALSE, include=FALSE}
library(tidyverse)
library(PPP)
library(parallel)
library(dplyr)
library(stats4)
library(clipp)
library(survival)


source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/prior_elicitation.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/helpers.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/outputHelpers.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mhLoglikehood_clipp.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mcmcPenetrance.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/InternalVars.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Additional Scripts/EstimEval.R")

set.seed(2024)
```

# Simulated Families

```{r desc fams, echo=FALSE, cache=TRUE}

# Function to describe the families
describeFamilies <- function(fams, cancer, gene) {
  # Validate and translate cancer type
  if (!cancer %in% CANCER_NAME_MAP$long) {
    stop("Cancer type '", cancer, "' is not supported. Please choose from the supported list.")
  }
  
  # Get the abbreviation for the cancer type
  cancer_index <- which(CANCER_NAME_MAP$long == cancer)
  cancer_short <- CANCER_NAME_MAP$short[cancer_index]
  
  # Construct column names based on cancer type and gene input
  aff_col_name <- paste0("isAff", cancer_short)
  age_col_name <- paste0("Age", cancer_short)
  
  # Initialize counters and vectors for storing data
  affectedFamilies = 0 
  affectedProbands = 0
  famSizes = c()
  geneFamilies = 0
  geneProbands = 0 
  curAges = c()
  cancerAges = c()
  pbCancerAges = c()
  pbCurAges = c()
  pbCancerAgesFemale = c()
  affectedFemaleProbands = 0
  geneFemaleProbands = 0 
  cancerAgesFemale = c()
  total_individuals = 0
  gene_individuals = 0
  
  # Process each family
  for(i in 1:length(fams)){
    f = fams[[i]]
    sizeOfFamily = nrow(f)
    total_individuals = total_individuals + sizeOfFamily
    famSizes = c(famSizes, sizeOfFamily)
    curAges = c(curAges, f$CurAge)  
    f = f %>% dplyr::filter(!!rlang::sym(aff_col_name) == 1, isProband == 0) 
    cancerAges <- c(cancerAges, f[[age_col_name]])  
    ff = fams[[i]] %>% dplyr::filter(!!rlang::sym(gene) == 1, !!rlang::sym(aff_col_name) == 1) 
    gene_individuals = gene_individuals + nrow(ff)
    cancerAgesFemale <- c(cancerAgesFemale, ff[[age_col_name]])  

    if(nrow(f)>0){
      affectedFamilies = affectedFamilies + 1
    }
    geneFams = fams[[i]] %>% dplyr::filter(isProband == 0, !!rlang::sym(gene) == 1)
    if(nrow(geneFams) > 0){
     geneFamilies = geneFamilies + 1
    }
    
    pb <- fams[[i]] %>% dplyr::filter(isProband == 1)
    pbfemale <- fams[[i]] %>% dplyr::filter(isProband == 1, Sex == 0)
    
    if (nrow(pbfemale) > 0) {
      pbCancerAgesFemale <- c(pbCancerAgesFemale, pbfemale[[age_col_name]])
      if (any(pbfemale[[aff_col_name]] == 1)){
        affectedFemaleProbands <- affectedFemaleProbands + 1
      }
      if (!is.na(pbfemale[[gene]])){
        if (any(pbfemale[[gene]] == 1)){
          geneFemaleProbands <- geneFemaleProbands + 1
        }
      }
    }
  
    pbCancerAges <- c(pbCancerAges, pb[[age_col_name]])
    pbCurAges <- c(pbCurAges, pb$CurAge)  
    
    if (any(pb[[aff_col_name]] == 1)){
      affectedProbands <- affectedProbands + 1
    }
    
    if (!is.na(pb[[gene]])){
      if (any(pb[[gene]] == 1)){
        geneProbands <- geneProbands + 1
      }
    }
  }

  print(paste0("Number of families: ", length(fams)))
  print(paste0("Total Individuals: ", total_individuals))
  print(paste0("Average family size: ", mean(famSizes)))
  print("Summary of family sizes")
  print(summary(famSizes))
  print("Summary of Current Age for all family members")
  print(summary(curAges))
  print("Summary of Current Age for probands")
  print(summary(pbCurAges))
  
  print(paste0("Number of families with affected individuals (", cancer, "): ", 
               affectedFamilies))
  print(paste0("Number of families with affected probands (", cancer, "): ", 
               affectedProbands))
  print(paste0("Number of families with affected female probands (", cancer, "): ", 
               affectedFemaleProbands))
  print(paste0("Number of families with relatives with PV (and not the proband) (", gene, "): ", geneFamilies))
  print(paste0("Number of probands with PV (", gene, "): ", geneProbands))
  print(paste0("Number of female probands with PV (", gene, "): ", 
               geneFemaleProbands))
  print(paste0("Number of individuals with PV (", gene, "): ", gene_individuals))
  
  print("Summary of Cancer Age of (female) probands")
  print(summary(pbCancerAgesFemale))

  print("Summary of Cancer Age in affected (non-proband) individuals")
  print(summary(cancerAges))
}

# Load the families of the simulation
# Load the additonal families
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Simulated\ Families/simfamilies_sim1.RData")
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Simulated\ Families/simfamilies_sim2.RData")
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Simulated\ Families/simfamilies_sim3.RData")
```

We simulated three different sizes of families. For each simulation run we subsetted the families to only include those families where the proband was a carrier of the PV. The simulation was done using the PedUtils package. The penetrance for non-carriers is based on the SEER penetrances in the PanelPRO Database. For carriers we specify a Weibull penetrance with the following parameters:

-   Alpha = 2.5

-   Beta = 50

-   Delta = 20

-   Gamma = 0.9

Censoring is based on a $$N(85,10)$$. The simulated age of death is centered at 85 with a standard deviation of 10, but values above ageMax (default = 94) are re-simulated to be centered around ageMax - 15.

## Family Sim 1

```{r fam1, echo=FALSE, cache=TRUE}
print(describeFamilies(simfamilies_sim1, "Breast","BRCA1"))
```

## Family Sim 2

```{r fam2, echo=FALSE, cache=TRUE}
print(describeFamilies(simfamilies_sim2, "Breast","BRCA1"))
```

## Family Sim 3

```{r fam3, echo=FALSE, cache=TRUE}
print(describeFamilies(simfamilies_sim3, "Breast","BRCA1"))
```

# Likelihood

The likelihood model in `penet.fn` is based on the Weibull distribution for carriers and, per default, the SEER peentrance data for non-carriers. For an individual $i$, the model considers the following parameters:

-   $\alpha$: Shape parameter of the Weibull distribution
-   $\beta$: Scale parameter of the Weibull distribution
-   $\delta$: Shift parameter for the age in the Weibull function
-   $\gamma$: Asymptote parameter to adjust the scale of the Weibull function
-   $\text{age}_i$: Age of the individual $i$
-   $\text{aff}_i$: Affection status (1 if affected, 0 otherwise)

An indiviudal that develops cancer at $\text{age}_i$ contributes a term $f$ to the likelihood, which is just the density of the Weibull in our model. The contribution from an individual whose survival time is censored at is the survival probability $S_i = 1 - F$ where $F$ is corresponding cumulative distribution function of the Weibull.

The survival probability for individual $i$ is calculated as: $$S_i = 1 - pweibull(\text{age}_i - \delta, \alpha, \beta) \times \gamma$$ where pweibull is the CDF of the Weibull.

The density (for affected individuals) is calculated as:

$$ c.pen_i = dweibull(\text{age}_i - \delta, \alpha, \beta) \times \gamma $$

For the function \( f(x; \alpha, \beta, \delta, \gamma) \) we define:

$$
f(x; \alpha, \beta, \delta, \gamma) = 
\begin{cases} 
\gamma \left( \frac{\alpha}{\beta} \left( \frac{x - \delta}{\beta} \right)^{\alpha - 1} e^{-\left( \frac{x - \delta}{\beta} \right)^\alpha } \right), & x \geq \delta \\
0, & x < \delta
\end{cases}
$$

And for the cumulative distribution function \( F(x; \alpha, \beta, \delta, \gamma) \):

$$
F(x; \alpha, \beta, \delta, \gamma) =
\begin{cases} 
\gamma \left( 1 - e^{-\left( \frac{x - \delta}{\beta} \right)^\alpha } \right), & x \geq \delta \\
0, & x < \delta
\end{cases}
$$


The baseline risk $\text{SEER\_baseline}_i$ is derived from SEER data:

$$
\text{nc.pen}_i = 
\begin{cases} 
\text{SEER\_baseline}_i & \text{if SeerNC is TRUE} \\
\text{Custom non-carrier penetrance model} & \text{otherwise}
\end{cases}
$$

For censored observations, the penetrance is considered as the cumulative probability of being cancer-free, while for affected observations, it is based on the calculated density at their age of diagnosis.

Hence, the penetrance $\text{penet}_i$ for individual $i$ and for all three potential genotypes 1/1 (non-carriers), 1/2 (heterozygous carrier), and 2/2 (homozygous carrier) is:

$$
\text{penet}_i = 
\begin{cases} 
1 - \sum_{j=1}^{\text{age}_i} \text{SEER\_baseline}_j, \text{Survival\_prob}_i, \text{Survival\_prob}_i & \text{if } \text{aff}_i = 0 \\
\text{nc.pen}_i, c.pen_i, c.pen_i & \text{if } \text{aff}_i = 1
\end{cases}
$$

The adjustments for observed genotypes, gender-specific handling, and the condition for homozygotes are given as follows.

\begin{verbatim}
if (data$geno[i] == "1/1") penet.i[-1] <- 0
if (data$geno[i] == "1/2") penet.i[-2] <- 0
if (data$geno[i] == "2/2") penet.i[-3] <- 0

if (!homozygote) {
    penet.i[3] <- 0
}

pen <- 1e-28
if (sex == "Male" && data$sex[i] != 1) penet.i <- c(pen, pen, pen)
if (sex == "Female" && data$sex[i] != 2) penet.i <- c(pen, pen, pen)
if (sex == "NA") penet.i <- penet.i

return(penet.i)
\end{verbatim}

\begin{align*}
\text{For } i &\in \{1, 2, \ldots, n\}: \\
&\text{if } \text{geno}_i = 1/1, \text{ then } \text{penet}_i[-1] = 0 \\
&\text{if } \text{geno}_i = 1/2, \text{ then } \text{penet}_i[-2] = 0 \\
&\text{if } \text{geno}_i = 2/2, \text{ then } \text{penet}_i[-3] = 0 \\
&\text{if not homozygote, then penet}_i[3] = 0 \\
&\text{Set pen} = 1 \times 10^{-28} \\
&\text{if sex is Male and } \text{sex}_i \neq 1, \text{ then penet}_i = \text{pen} \\
&\text{if sex is Female and } \text{sex}_i \neq 2, \text{ then penet}_i = \text{pen} \\
&\text{if sex is NA, then } \text{penet}_i \text{ remains unchanged}
\end{align*}

Based on the above a penetrance matrix is constructed for the pedigree. The `(i, k)`^th^ component `penet[i, k]` of the penetrance matrix is the conditional probability of person `i`'s observed data $x_i$, given that his or her actual genotype $g_i$ is the `k`^th^ genotype. Given a genetic model and a penetrance matrix, we then use `clipp` to calculate the loglikelihood for all pedigrees.

# Families Sim1

## Default Prior

In this simulation the penetrance estimation is applied to the first set of families from sim1. The default prior settings are applied here with a uniform prior on the asymptote parameter.

```{r, echo=TRUE, cache=TRUE}
dat <- simfamilies_sim1

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 1, g2 = 1),
    threshold = list(min = 15, max = 25),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
out_sim_1 <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 100, 
    prior_params = prior_params_BRCA1, af = 0.1, burn_in = 0.1, sex = "Female"
)
save(out_sim_1, file = "out_sim_1.Rdata")

# print summary stats
out_sim_1$summary_stats
plot_traceSingle(out_sim_1$results[[1]])
ages <- 1:94
plot_penetrance_sim(out_sim_1$combined_chains, prob = 0.95, max_age = 94,2.5,50,20,0.9)

# Plot the running mean 

runFQ<- running_mean(out_sim_1$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_1$results[[1]]$median_samples)
runThreshold <- running_mean(out_sim_1$results[[1]]$threshold_samples)
runAsy <- running_mean(out_sim_1$results[[1]]$asymptote_samples)

# Plot running mean
#plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
#plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
#plot(runThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Threshold Samples")
#plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_1$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_1$results[[1]]$median_samples)
runVThreshold <- running_variance(out_sim_1$results[[1]]$threshold_samples)
runVAsy <- running_variance(out_sim_1$results[[1]]$asymptote_samples)

# Plot running var
#plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
#plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
#plot(runVThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Threshold Samples")
#plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")


  # Autocorrelation plots
acf(out_sim_1$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_1$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_1$results[[1]]$threshold_samples, main = "ACF of Threshold")
acf(out_sim_1$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

# plot the scatterplot with the median and first quartile
plot(out_sim_1$combined_chains$median_results, out_sim_1$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```

## Fixed Asymptote

Now we rerun the simulation but with the prior narrowly centered around the known asymptote parameter of $\gamma = 0.9$.

```{r, echo=TRUE, cache=TRUE}

# Load the additonal families
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Simulated\ Families/simfamilies_sim1.RData")
dat <- simfamilies_sim1

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 900, g2 = 100),
    threshold = list(min = 15, max = 25),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
out_sim_1_fixed <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 100, 
    prior_params = prior_params_BRCA1, af = 0.1, burn_in = 0.1, sex = "Female"
)
save(out_sim_1_fixed, file = "out_sim_1_fixed.Rdata")

# print summary stats
out_sim_1_fixed$summary_stats
plot_traceSingle(out_sim_1_fixed$results[[1]])
ages <- 1:94
plot_penetrance_sim(out_sim_1_fixed$combined_chains, prob = 0.95, max_age = 94,2.5,50,20,0.9)

# Plot the running mean 

runFQ<- running_mean(out_sim_1_fixed$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_1_fixed$results[[1]]$median_samples)
runThreshold <- running_mean(out_sim_1_fixed$results[[1]]$threshold_samples)
runAsy <- running_mean(out_sim_1_fixed$results[[1]]$asymptote_samples)

# Plot running mean
#plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
#plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
#plot(runThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Threshold Samples")
#plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_1_fixed$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_1_fixed$results[[1]]$median_samples)
runVThreshold <- running_variance(out_sim_1_fixed$results[[1]]$threshold_samples)
runVAsy <- running_variance(out_sim_1_fixed$results[[1]]$asymptote_samples)

# Plot running var
#plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
#plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
#plot(runVThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Threshold Samples")
#plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")

  # Autocorrelation plots
acf(out_sim_1_fixed$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_1_fixed$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_1_fixed$results[[1]]$threshold_samples, main = "ACF of Threshold")
acf(out_sim_1_fixed$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

# plot the scatterplot with the median and first quartile
plot(out_sim_1_fixed$combined_chains$median_results, out_sim_1_fixed$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```

# Families Sim2

## Default Prior

```{r, echo=TRUE, cache=TRUE}
dat <- simfamilies_sim2

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 1, g2 = 1),
    threshold = list(min = 15, max = 25),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
out_sim_2 <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 100, 
    prior_params = prior_params_BRCA1, af = 0.1, burn_in = 0.1, sex = "Female"
)
save(out_sim_2, file = "out_sim_2.Rdata")

# print summary stats
out_sim_2$summary_stats
plot_traceSingle(out_sim_2$results[[1]])
plot_penetrance_sim(out_sim_2$combined_chains, prob = 0.95, max_age = 94,2.5,50,20,0.9)

# Plot the running mean 

runFQ<- running_mean(out_sim_2$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_2$results[[1]]$median_samples)
runThreshold <- running_mean(out_sim_2$results[[1]]$threshold_samples)
runAsy <- running_mean(out_sim_2$results[[1]]$asymptote_samples)

# Plot running mean
#plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
#plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
#plot(runThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Threshold Samples")
#plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_2$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_2$results[[1]]$median_samples)
runVThreshold <- running_variance(out_sim_2$results[[1]]$threshold_samples)
runVAsy <- running_variance(out_sim_2$results[[1]]$asymptote_samples)

# Plot running var
#plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
#plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
#plot(runVThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Threshold Samples")
#plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")

# Autocorrelation plots
acf(out_sim_2$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_2$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_2$results[[1]]$threshold_samples, main = "ACF of Threshold")
acf(out_sim_2$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

# plot the scatterplot with the median and first quartile
plot(out_sim_2$combined_chains$median_results, out_sim_2$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```

# Families Sim3

## Default Prior

```{r, echo=TRUE, cache=TRUE}

# Load the additonal families
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Simulated\ Families/simfamilies_sim3.RData")
dat <- simfamilies_sim3

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 1, g2 = 1),
    threshold = list(min = 15, max = 25),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
out_sim_3 <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 200, 
    prior_params = prior_params_BRCA1, af = 0.1, burn_in = 0.1, sex = "Female"
)
save(out_sim_3, file = "out_sim_3.Rdata")

# print summary stats
out_sim_3$summary_stats
plot_traceSingle(out_sim_3$results[[1]])
ages <- 1:94
plot_penetrance_sim(out_sim_3$combined_chains, prob = 0.95, max_age = 94,"Female",2.5,50,20,0.1,0.9)

# Plot the running mean 

runFQ<- running_mean(out_sim_3$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_3$results[[1]]$median_samples)
runThreshold <- running_mean(out_sim_3$results[[1]]$threshold_samples)
runAsy <- running_mean(out_sim_3$results[[1]]$asymptote_samples)

# Plot running mean
#plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
#plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
#plot(runThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Threshold Samples")
#plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")

# Plot the running var 
runVFQ<- running_variance(out_sim_3$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_3$results[[1]]$median_samples)
runVThreshold <- running_variance(out_sim_3$results[[1]]$threshold_samples)
runVAsy <- running_variance(out_sim_3$results[[1]]$asymptote_samples)

# Plot running var
#plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
#plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
#plot(runVThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Threshold Samples")
#plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")

# Autocorrelation plots
acf(out_sim_3$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_3$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_3$results[[1]]$threshold_samples, main = "ACF of Threshold")
acf(out_sim_3$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

# plot the scatterplot with the median and first quartile
plot(out_sim_3$combined_chains$median_results, out_sim_3$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```

## Fixed Prior

```{r, echo=TRUE, cache=TRUE}

# Load the additonal families
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Simulated\ Families/simfamilies_sim3.RData")
dat <- simfamilies_sim3

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 900, g2 = 100),
    threshold = list(min = 15, max = 25),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
out_sim_3_fixed<- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 100, 
    prior_params = prior_params_BRCA1, af = 0.1, burn_in = 0.1, sex = "Female"
)
save(out_sim_3_fixed, file = "out_sim_3_fixed.Rdata")

# print summary stats
out_sim_3_fixed$summary_stats
plot_traceSingle(out_sim_3_fixed$results[[1]])
ages <- 1:94
plot_penetrance_sim(out_sim_3_fixed$combined_chains, prob = 0.95, max_age = 94,2.5,50,20,0.9)

# Plot the running mean 
runFQ<- running_mean(out_sim_3_fixed$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_3_fixed$results[[1]]$median_samples)
runThreshold <- running_mean(out_sim_3_fixed$results[[1]]$threshold_samples)
runAsy <- running_mean(out_sim_3_fixed$results[[1]]$asymptote_samples)

# Plot running mean
#plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
#plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
#plot(runThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Threshold Samples")
#plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")

# Plot the running var 
runVFQ<- running_variance(out_sim_3_fixed$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_3_fixed$results[[1]]$median_samples)
runVThreshold <- running_variance(out_sim_3_fixed$results[[1]]$threshold_samples)
runVAsy <- running_variance(out_sim_3_fixed$results[[1]]$asymptote_samples)

# Plot running var
#plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
#plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
#plot(runVThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Threshold Samples")
#plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")

# Autocorrelation plots
acf(out_sim_3_fixed$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_3_fixed$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_3_fixed$results[[1]]$threshold_samples, main = "ACF of Threshold")
acf(out_sim_3_fixed$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

# plot the scatterplot with the median and first quartile
plot(out_sim_3_fixed$combined_chains$median_results, out_sim_3_fixed$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```
