---
title: "PenEstim"
subtitle: "Simulation Study"
format: 
  html:
    toc: true
    toc-location: body
date: 3-19-2024
theme: lightly
author:
  - name: "Nicolas Kubista"
    affiliations: "Harvard University"
embed-resources: true
toc: true
toc-expand: 2
toc-title: Contents
page-layout: full
code-fold: true
editor: visual
---

```{r, echo=FALSE, include=FALSE}
library(tidyverse)
library(PPP)
library(parallel)
library(dplyr)
library(stats4)
library(clipp)
library(survival)


source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/prior_elicitation.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/helpers.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/outputHelpers.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mhLoglikehood_clipp.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mcmcPenetrance.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/InternalVars.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Additional Scripts/EstimEval.R")

set.seed(2024)
```

# Simulated Families

# Likelihood

The likelihood model in `penet.fn` is based on the Weibull distribution for carriers and, per default, the SEER peentrance data for non-carriers. For an individual $i$, the model considers the following parameters:

- $\alpha$: Shape parameter of the Weibull distribution
- $\beta$: Scale parameter of the Weibull distribution
- $\delta$: Shift parameter for the age in the Weibull function
- $\gamma$: Multiplier to adjust the scale of the Weibull function
- $\text{age}_i$: Age of the individual $i$
- $\text{aff}_i$: Affection status (1 if affected, 0 otherwise)

The survival probability for individual $i$ is given by:

$$\text{Survival\_prob}_i = 1 - pweibull(\text{age}_i - \delta, \alpha, \beta) \times \gamma$$
where pweibull is the CDF of the Weibull. 

The density (for affected individuals) is calculated as:

$$c.pen_i = dweibull(\text{age}_i - \delta, \alpha, \beta) \times \gamma$$

The baseline risk $\text{SEER\_baseline}_i$ is derived from SEER data:

$$
\text{nc.pen}_i = 
\begin{cases} 
\text{SEER\_baseline}_i & \text{if SeerNC is TRUE} \\
\text{Custom non-carrier penetrance model} & \text{otherwise}
\end{cases}
$$

For censored observations, the penetrance is considered as the cumulative probability of being cancer-free, while for affected observations, it is based on the calculated density at their age of diagnosis.

Hence, the penetrance $\text{penet}_i$ for individual $i$ and for all three potential genotypes 1/1 (non-carriers), 1/2 (heterozygous carrier), and 2/2 (homozygous carrier) is:

$$
\text{penet}_i = 
\begin{cases} 
1 - \sum_{j=1}^{\text{age}_i} \text{SEER\_baseline}_j, \text{Survival\_prob}_i, \text{Survival\_prob}_i & \text{if } \text{aff}_i = 0 \\
\text{nc.pen}_i, c.pen_i, c.pen_i & \text{if } \text{aff}_i = 1
\end{cases}
$$

# Families Sim1

## Default Prior

In this simulation the penetrance estimation is applied to the first set of families from sim1. The default prior settings are applied here with a uniform prior on the asymptote parameter.

```{r, echo=TRUE, cache=TRUE}

# Load the additonal families
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Simulated\ Families/simfamilies_sim1.RData")
dat <- simfamilies_sim1

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 1, g2 = 1),
    threshold = list(min = 15, max = 25),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
out_sim_1 <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 100, 
    prior_params = prior_params_BRCA1, af = 0.1, burn_in = 0.1, sex = "Female"
)
save(out_sim_1, file = "out_sim_1.Rdata")

# print summary stats
out_sim_1$summary_stats
plot_traceSingle(out_sim_1$results[[1]])
ages <- 1:94
plot_penetrance_sim(out_sim_1$combined_chains, prob = 0.95, max_age = 94,2.5,50,20,0.9)

# Plot the running mean 

runFQ<- running_mean(out_sim_1$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_1$results[[1]]$median_samples)
runThreshold <- running_mean(out_sim_1$results[[1]]$threshold_samples)
runAsy <- running_mean(out_sim_1$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Threshold Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_1$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_1$results[[1]]$median_samples)
runVThreshold <- running_variance(out_sim_1$results[[1]]$threshold_samples)
runVAsy <- running_variance(out_sim_1$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Threshold Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")


  # Autocorrelation plots
acf(out_sim_1$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_1$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_1$results[[1]]$threshold_samples, main = "ACF of Threshold")
acf(out_sim_1$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

# plot the scatterplot with the median and first quartile
plot(out_sim_1$combined_chains$median_results, out_sim_1$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```

## Fixed Asymptote

Now we rerun the simulation but with the prior narrowly centered around the known asymptote parameter of $\gamma = 0.9$.

```{r, echo=TRUE, cache=TRUE}

# Load the additonal families
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Simulated\ Families/simfamilies_sim1.RData")
dat <- simfamilies_sim1

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 900, g2 = 100),
    threshold = list(min = 15, max = 25),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
out_sim_1_fixed <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 10, 
    prior_params = prior_params_BRCA1, af = 0.1, burn_in = 0.1, sex = "Female"
)
save(out_sim_1_fixed, file = "out_sim_1_fixed.Rdata")

# print summary stats
out_sim_1_fixed$summary_stats
plot_traceSingle(out_sim_1_fixed$results[[1]])
ages <- 1:94
plot_penetrance_sim(out_sim_1_fixed$combined_chains, prob = 0.95, max_age = 94,2.5,50,20,0.9)

# Plot the running mean 

runFQ<- running_mean(out_sim_1_fixed$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_1_fixed$results[[1]]$median_samples)
runThreshold <- running_mean(out_sim_1_fixed$results[[1]]$threshold_samples)
runAsy <- running_mean(out_sim_1_fixed$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Threshold Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_1_fixed$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_1_fixed$results[[1]]$median_samples)
runVThreshold <- running_variance(out_sim_1_fixed$results[[1]]$threshold_samples)
runVAsy <- running_variance(out_sim_1_fixed$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Threshold Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")


  # Autocorrelation plots
acf(out_sim_1_fixed$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_1_fixed$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_1_fixed$results[[1]]$threshold_samples, main = "ACF of Threshold")
acf(out_sim_1_fixed$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

# plot the scatterplot with the median and first quartile
plot(out_sim_1_fixed$combined_chains$median_results, out_sim_1_fixed$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```

# Families Sim2

## Default Prior

```{r, echo=TRUE, cache=TRUE}

# Load the additonal families
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Simulated\ Families/simfamilies_sim2.RData")
dat <- simfamilies_sim2

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 1, g2 = 1),
    threshold = list(min = 15, max = 25),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
out_sim_2 <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 111, 
    prior_params = prior_params_BRCA1, af = 0.1, burn_in = 0.1, sex = "Female"
)
save(out_sim_2, file = "out_sim_2.Rdata")

# print summary stats
out_sim_2$summary_stats
plot_traceSingle(out_sim_2$results[[1]])
plot_penetrance_sim(out_sim_2$combined_chains, prob = 0.95, max_age = 94,2.5,50,20,0.9)

# Plot the running mean 

runFQ<- running_mean(out_sim_2$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_2$results[[1]]$median_samples)
runThreshold <- running_mean(out_sim_2$results[[1]]$threshold_samples)
runAsy <- running_mean(out_sim_2$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Threshold Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_2$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_2$results[[1]]$median_samples)
runVThreshold <- running_variance(out_sim_2$results[[1]]$threshold_samples)
runVAsy <- running_variance(out_sim_2$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Threshold Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")


  # Autocorrelation plots
acf(out_sim_2$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_2$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_2$results[[1]]$threshold_samples, main = "ACF of Threshold")
acf(out_sim_2$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

# plot the scatterplot with the median and first quartile
plot(out_sim_2$combined_chains$median_results, out_sim_2$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```

# Families Sim3

## Default Prior

```{r, echo=TRUE, cache=TRUE}

# Load the additonal families
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Simulated\ Families/simfamilies_sim3.RData")
dat <- simfamilies_sim3

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 1, g2 = 1),
    threshold = list(min = 15, max = 25),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
out_sim_3 <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 11, 
    prior_params = prior_params_BRCA1, af = 0.1, burn_in = 0.1, sex = "Female"
)
save(out_sim_3, file = "out_sim_3.Rdata")

# print summary stats
out_sim_3$summary_stats
plot_traceSingle(out_sim_3$results[[1]])
ages <- 1:94
plot_penetrance_sim(out_sim_3$combined_chains, prob = 0.95, max_age = 94,2.5,50,20,0.9)

# Plot the running mean 

runFQ<- running_mean(out_sim_3$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_3$results[[1]]$median_samples)
runThreshold <- running_mean(out_sim_3$results[[1]]$threshold_samples)
runAsy <- running_mean(out_sim_3$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Threshold Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_3$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_3$results[[1]]$median_samples)
runVThreshold <- running_variance(out_sim_3$results[[1]]$threshold_samples)
runVAsy <- running_variance(out_sim_3$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Threshold Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")


  # Autocorrelation plots
acf(out_sim_3$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_3$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_3$results[[1]]$threshold_samples, main = "ACF of Threshold")
acf(out_sim_3$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

# plot the scatterplot with the median and first quartile
plot(out_sim_3$combined_chains$median_results, out_sim_3$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```

## Fixed Prior

```{r, echo=TRUE, cache=TRUE}

# Load the additonal families
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Simulated\ Families/simfamilies_sim3.RData")
dat <- simfamilies_sim3

# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 900, g2 = 100),
    threshold = list(min = 15, max = 25),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
out_sim_3_fixed<- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 111, 
    prior_params = prior_params_BRCA1, af = 0.1, burn_in = 0.1, sex = "Female"
)
save(out_sim_3_fixed, file = "out_sim_3_fixed.Rdata")

# print summary stats
out_sim_3_fixed$summary_stats
plot_traceSingle(out_sim_3_fixed$results[[1]])
ages <- 1:94
plot_penetrance_sim(out_sim_3_fixed$combined_chains, prob = 0.95, max_age = 94,2.5,50,20,0.9)

# Plot the running mean 

runFQ<- running_mean(out_sim_3_fixed$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_3_fixed$results[[1]]$median_samples)
runThreshold <- running_mean(out_sim_3_fixed$results[[1]]$threshold_samples)
runAsy <- running_mean(out_sim_3_fixed$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Threshold Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_3_fixed$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_3_fixed$results[[1]]$median_samples)
runVThreshold <- running_variance(out_sim_3_fixed$results[[1]]$threshold_samples)
runVAsy <- running_variance(out_sim_3_fixed$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVThreshold, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Threshold Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")


  # Autocorrelation plots
acf(out_sim_3_fixed$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_3_fixed$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_3_fixed$results[[1]]$threshold_samples, main = "ACF of Threshold")
acf(out_sim_3_fixed$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

# plot the scatterplot with the median and first quartile
plot(out_sim_3_fixed$combined_chains$median_results, out_sim_3_fixed$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```
