---
title: "PenEstim Application"
format:
  html:
    theme: default
---

```{r, echo=FALSE, include=FALSE}
library(tidyverse)
library(PPP)
library(parallel)
library(dplyr)
library(stats4)
library(clipp)
library(survival)


source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/prior_elicitation.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/helpers.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/outputHelpers.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mhLoglikelihood.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mhLoglikehood_clipp.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mcmcPenetrance.R")
```

**Data:**

We use simulated data and plot the penetrance using Kaplan-Mayer estimator. The families were simulated using the PedUtils package. It contains 500 families with 19 carriers of the PV.

```{r desc, echo=FALSE}

# Simulation study
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Simulated Families/families.RData")
dat <- families

describeFamilies <- function(fams){
  
  affectedFamilies = 0 
  affectedProbands = 0
  famSizes = c()
  BRCA1Families = 0
  BRCA1Probands = 0 
  curAges = 0
  BCAges = 0
  pbBCAges=0
  pbCurAges=0
  pbBCAgesfemale = 0
  affectedfemaleProbands = 0
  BRCA1femaleProbands = 0 
  BCAgesfemale = 0
  total_individuals = 0
  BRCA1_individuals = 0

  #affectedFamiliesList = list()
  #j = 1
  for(i in 1:length(fams)){
    f = fams[[i]]
    sizeOfFamily = nrow(f)
    total_individuals = total_individuals + sizeOfFamily
    famSizes = c(famSizes, sizeOfFamily)
    curAges = c(curAges, f$CurAge)  
    f = f %>% filter(isAffBC==1) %>% filter(isProband==0) 
    BCAges <- c(BCAges, f$AgeBC)  
    ff = fams[[i]] %>% filter(BRCA1==1) %>% filter(isAffBC==1) 
    BRCA1_individuals = BRCA1_individuals + nrow(ff)
    BCAgesfemale <- c(BCAgesfemale, ff$AgeBC)  

    if(nrow(f)>0){
      affectedFamilies = affectedFamilies + 1
    
    }
    BRCA1fams = fams[[i]] %>% filter(isProband==0) %>% filter(BRCA1==1)
    if(nrow(BRCA1fams) > 0){
     BRCA1Families = BRCA1Families + 1
    }
    
    pb <- fams[[i]] %>% filter(isProband == 1)
    pbfemale <- fams[[i]] %>% filter(isProband == 1) %>% filter(Sex == 0)
    
    
    # Check if pbfemale has any rows before accessing its columns
    if (nrow(pbfemale) > 0) {
      pbBCAgesfemale <- c(pbBCAgesfemale, pbfemale$AgeBC)
      if (any(pbfemale$isAffBC == 1)){
        affectedfemaleProbands <- affectedfemaleProbands + 1
      }
      if (!is.na(pbfemale$BRCA1)){
        if (any(pbfemale$BRCA1 == 1)){
          BRCA1femaleProbands <- BRCA1femaleProbands + 1
        }
      }
    }
  
    pbBCAges <- c(pbBCAges, pb$AgeBC)
    pbCurAges <- c(pbCurAges, pb$CurAge)  
    
    if (any(pb$isAffBC == 1)){
      affectedProbands <- affectedProbands + 1
    }
    
        if (!is.na(pb$BRCA1)){
      if (any(pb$BRCA1 == 1)){
        BRCA1Probands <- BRCA1Probands + 1
      }
    }
  }
    
 
  
  # careful with the description 
  # do the stratified descriptives 
  # check the weibull to look like 
  
  print(paste0("Number of families: ", length(fams)))
  print(paste0("Total Individuals: ", total_individuals))
  print(paste0("Average family size: ", mean(famSizes)))
  print("Summary of family sizes")
  print(summary(famSizes))
  print("Summary of Current Age for all family members")
  print(summary(curAges))
  print("Summary of Current Age for probands")
  print(summary(pbCurAges))
  
  print(paste0("Number of families with affected (BC) individuals: ", 
               affectedFamilies))
  print(paste0("Number of families with affected (BC) probands: ", 
               affectedProbands))
  print(paste0("Number of families with affected (BC) female probands: ", 
               affectedfemaleProbands))
  print(paste0("Number of families with relatives with PV (and not the proband)
               : ", BRCA1Families))
  print(paste0("Number of families with proband with PV: ", BRCA1Probands))
  print(paste0("Number of families with female proband with PV: ", 
               BRCA1femaleProbands))
  print(paste0("Number of individuals with PV: ", BRCA1_individuals))
  
  print("Summary of BC Age of (female) probands")
  print(summary(pbBCAgesfemale))

  print("Summary of BC Age in affected (non-proband) individuals")
  print(summary(BCAges))
  
  print("Summary of BC Age in affected (non-proband) female individuals")
  print(summary(BCAgesfemale))
}

print(describeFamilies(dat))


```

**Pentrance Curve using Kaplan-Meyer Estimator**

```{r, echo=FALSE}


allcarriers <- data.frame()
allnoncarriers <- data.frame()

# Plot the empirical penetrance curve
plot_penetrance_curves <- function(fams) {
  # Create an empty plot
  plot(0, 0, type = "n", xlim = c(0, 100), ylim = c(0, 1), xlab = "Age", 
       ylab = "Penetrance",
       main = "Estimated Penetrance Curves (for females)")

  # Initialize colors and line types
  colors <- c("red", "blue")  
  line_types <- c(1, 2)  

  # Create empty data frames to aggregate the data
  allcarriers <- data.frame()
  allnoncarriers <- data.frame()

  # Iterate through each family and aggregate the data
  for (i in 1:length(fams)) {
    f <- fams[[i]]
    carriers <- f %>% filter(BRCA1 == 1) %>% filter(Sex == 0)
    noncarriers <- f %>% filter(BRCA1 == 0) %>% filter(Sex == 0)
    
    # Append the data to the combined data frames
    allcarriers<- rbind(allcarriers, carriers)
    allnoncarriers <- rbind(allnoncarriers, noncarriers)
  }
  
  # Convert the cancer status variable to survival objects for each genotype
  surv_obj_1 <- Surv(time = allcarriers$AgeBC, event = allcarriers$isAffBC)
  surv_obj_0 <- Surv(time = allnoncarriers$AgeBC, event = allnoncarriers$isAffBC)
  
  # Estimate the Kaplan-Meier survival curves for each genotype
  km_fit_1 <- survfit(surv_obj_1 ~ 1, data = allcarriers)
  km_fit_0 <- survfit(surv_obj_0 ~ 1, data = allnoncarriers)
  
  # Calculate the penetrance functions for carriers and non-carriers
  penetrance_1 <- 1 - km_fit_1$surv
  penetrance_0 <- 1 - km_fit_0$surv
  
  # Plot the aggregated penetrance curves
  lines(km_fit_1$time, penetrance_1, type = "l", col = colors[1], 
        lty = line_types[1])
  lines(km_fit_0$time, penetrance_0, type = "l", col = colors[2], 
        lty = line_types[2])
  
  # Add legend
  legend("topright", legend = c("PV", "Non-PV"), col = colors[1:2], lty = 1,
         cex = 0.8)
}


# Plot the empirical penetrance curves
plot_penetrance_curves(dat)



```

**Estimation using PenEstim**

Application of the PenEstim algorithm using the clipp package for the likelihood calculation and the default prior settings.

-   First Quartile \$Q\_{25}\$ & Scaled Beta(6,3)

-   Median \$Q\_{50}\$ & Scaled Beta(2,2)

-   Asymptote \$\gamma\$ & Scaled Beta(1,1)

-   Threshold \$\delta\$ & Uniform(0,25)

```{r, echo=TRUE}

# Data Prep for the clipp structure from the classic PP format 
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "PedigreeID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}


# Run Estimation procedure with default prior setting 
out_sim_100 <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 100
)

# print summary stats
out_sim_100$summary_stats
```

**Review: Weibull parameter estimation**

The quantile function for the Weibull distribution, adjusted by a shift, is defined as:

$Q(p; \beta, \alpha, \delta) = \delta + \beta \left(-\log(1 - p)\right)^{\frac{1}{\alpha}}]$

where:

\- $p$ is the quantile probability,

\- $\beta$ is the scale parameter of the distribution,

\- $\alpha$ is the shape parameter of the distribution,

\- $\delta$ is a constant offset value.

The conditional median $Q_{50} = Q(0.5)$ and the conditional first quartile $Q_{25} = Q(0.25)$ are then:

$Q_{50} = \delta + \beta [ln (2)]^{1/\alpha}$

$Q_{25} = \delta + \beta[ln(4 / 3)]^{1/\alpha}$

The objective function, which is minimized to calculate the Weibull parameters given the median and first quartile drawn as from the prior distributions, is given by:

$f(\beta, \alpha) = (\frac{\text{given\_median}}{\text{asymptote}} - Q_{50})^2 + (\frac{\text{given\_first\_quartile}}{\text{asymptote}} - Q_{25})^2$

We then use an optimization process that seeks the values of $\alpha$ and $\beta$ that minimize this objective function.

Analytically, a solution is more challenging:

\*\*Step 1: Express $\beta$ in terms of $Q\_{50}$, $\delta$, and $\alpha$\*\*

From the equation for $Q\_{50}$, we can express $\beta$ as:

$\beta = \frac{Q\_{50} - \delta}{[\ln(2)]^{1/\alpha}}$

\*\*Step 2: Substitute $\beta$ into the equation for $Q\_{25}$\*\*

Substitute the expression for $\beta$ from $Q\_{50}$ into the equation for $Q\_{25}$:

$Q\_{25} = \delta + (\frac{Q\_{50} - \delta}{[ln(2)]^{1/\alpha}})[ln(4 / 3)]^{1/\alpha}$

Now solving for $\alpha$ is challenging...
