---
title: "PenEstim Application"
format: 
  html:
    toc: true
    toc_float: true
    number_sections: true
    theme: cosmo
    left:
        - text: "Home"
          href: index.html
        - text: "Section 1"
          href: #section-1
        - text: "Section 2"
          href: #section-2
---

```{r, echo=FALSE, include=FALSE}
library(tidyverse)
library(PPP)
library(parallel)
library(dplyr)
library(stats4)
library(clipp)
library(survival)


source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/prior_elicitation.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/helpers.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/outputHelpers.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mhLoglikelihood.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mhLoglikehood_clipp.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/mcmcPenetrance.R")
source("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/PenEstim/R/InternalVars.R")

set.seed(2024)
```

# Simulation 

## Data

We use simulated data for the first application. Families C were simulated with a wider Weibull distribution with shape parameter, $\alpha = 2.5$, scale parameter, $\beta=50$, a shift of $\delta=20$, and an asymptote parameter $\gamma = 0.9$. This translates into an age-specific penetrance curve which does not allow for cancer age (AgeBC) to be lower than 20 with a median age of $Q_{50}=62$ and a first quartile of $Q_{25}=51$. The allele frequency was set at $0.9$. Only 60 families of this type were generated to improve run-time for the simulation study using the MCMC approach.

```{r desc, echo=FALSE, cache=TRUE}

# Simulation study
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/Submission/Simulated Families/simFamilies_C_1000_nocen_selected.RData")
dat <- simFamilies_C_1000_nocen_selected

describeFamilies <- function(fams, cancer, gene) {
  # Validate and translate cancer type
  if (!cancer %in% CANCER_NAME_MAP$long) {
    stop("Cancer type '", cancer, "' is not supported. Please choose from the supported list.")
  }
  
  # Get the abbreviation for the cancer type
  cancer_index <- which(CANCER_NAME_MAP$long == cancer)
  cancer_short <- CANCER_NAME_MAP$short[cancer_index]
  
  # Construct column names based on cancer type and gene input
  aff_col_name <- paste0("isAff", cancer_short)
  age_col_name <- paste0("Age", cancer_short)
  
  # Initialize counters and vectors for storing data
  affectedFamilies = 0 
  affectedProbands = 0
  famSizes = c()
  geneFamilies = 0
  geneProbands = 0 
  curAges = c()
  cancerAges = c()
  pbCancerAges = c()
  pbCurAges = c()
  pbCancerAgesFemale = c()
  affectedFemaleProbands = 0
  geneFemaleProbands = 0 
  cancerAgesFemale = c()
  total_individuals = 0
  gene_individuals = 0
  
  # Process each family
  for(i in 1:length(fams)){
    f = fams[[i]]
    sizeOfFamily = nrow(f)
    total_individuals = total_individuals + sizeOfFamily
    famSizes = c(famSizes, sizeOfFamily)
    curAges = c(curAges, f$CurAge)  
    f = f %>% dplyr::filter(!!rlang::sym(aff_col_name) == 1, isProband == 0) 
    cancerAges <- c(cancerAges, f[[age_col_name]])  
    ff = fams[[i]] %>% dplyr::filter(!!rlang::sym(gene) == 1, !!rlang::sym(aff_col_name) == 1) 
    gene_individuals = gene_individuals + nrow(ff)
    cancerAgesFemale <- c(cancerAgesFemale, ff[[age_col_name]])  

    if(nrow(f)>0){
      affectedFamilies = affectedFamilies + 1
    }
    geneFams = fams[[i]] %>% dplyr::filter(isProband == 0, !!rlang::sym(gene) == 1)
    if(nrow(geneFams) > 0){
     geneFamilies = geneFamilies + 1
    }
    
    pb <- fams[[i]] %>% dplyr::filter(isProband == 1)
    pbfemale <- fams[[i]] %>% dplyr::filter(isProband == 1, Sex == 0)
    
    if (nrow(pbfemale) > 0) {
      pbCancerAgesFemale <- c(pbCancerAgesFemale, pbfemale[[age_col_name]])
      if (any(pbfemale[[aff_col_name]] == 1)){
        affectedFemaleProbands <- affectedFemaleProbands + 1
      }
      if (!is.na(pbfemale[[gene]])){
        if (any(pbfemale[[gene]] == 1)){
          geneFemaleProbands <- geneFemaleProbands + 1
        }
      }
    }
  
    pbCancerAges <- c(pbCancerAges, pb[[age_col_name]])
    pbCurAges <- c(pbCurAges, pb$CurAge)  
    
    if (any(pb[[aff_col_name]] == 1)){
      affectedProbands <- affectedProbands + 1
    }
    
    if (!is.na(pb[[gene]])){
      if (any(pb[[gene]] == 1)){
        geneProbands <- geneProbands + 1
      }
    }
  }

  print(paste0("Number of families: ", length(fams)))
  print(paste0("Total Individuals: ", total_individuals))
  print(paste0("Average family size: ", mean(famSizes)))
  print("Summary of family sizes")
  print(summary(famSizes))
  print("Summary of Current Age for all family members")
  print(summary(curAges))
  print("Summary of Current Age for probands")
  print(summary(pbCurAges))
  
  print(paste0("Number of families with affected individuals (", cancer, "): ", 
               affectedFamilies))
  print(paste0("Number of families with affected probands (", cancer, "): ", 
               affectedProbands))
  print(paste0("Number of families with affected female probands (", cancer, "): ", 
               affectedFemaleProbands))
  print(paste0("Number of families with relatives with PV (and not the proband) (", gene, "): ", geneFamilies))
  print(paste0("Number of probands with PV (", gene, "): ", geneProbands))
  print(paste0("Number of female probands with PV (", gene, "): ", 
               geneFemaleProbands))
  print(paste0("Number of individuals with PV (", gene, "): ", gene_individuals))
  
  print("Summary of Cancer Age of (female) probands")
  print(summary(pbCancerAgesFemale))

  print("Summary of Cancer Age in affected (non-proband) individuals")
  print(summary(cancerAges))
  
  print("Summary of Cancer Age in affected (non-proband) female individuals")
  print(summary(cancerAgesFemale))
}

print(describeFamilies(dat,"Breast","BRCA1"))
```

 ## Pentrance Curve using Kaplan-Meyer Estimator

```{r, echo=FALSE, cache=TRUE}

allcarriers <- data.frame()
allnoncarriers <- data.frame()

# Function for plotting the penetrance curve using KM
plot_penetrance_curves <- function(fams) {
  # Initialize data frames
  allcarriers <- data.frame()
  allnoncarriers <- data.frame()

  # Create an empty plot
  plot(0, 0, type = "n", xlim = c(0, 100), ylim = c(0, 1), xlab = "Age", 
       ylab = "Penetrance", main = "Estimated Penetrance Curves (for females)")

  # Colors and line types
  colors <- c("red", "blue")  
  line_types <- c(1, 2)  

  # Aggregate data from each family
  for (i in 1:length(fams)) {
    f <- fams[[i]]
    carriers <- f %>% filter(BRCA1 == 1) %>% filter(Sex == 0)
    noncarriers <- f %>% filter(BRCA1 == 0) %>% filter(Sex == 0)
    
    allcarriers <- rbind(allcarriers, carriers)
    allnoncarriers <- rbind(allnoncarriers, noncarriers)
  }
 
  # Kaplan-Meier survival analysis
  surv_obj_1 <- Surv(time = allcarriers$AgeBC, event = allcarriers$isAffBC)
  surv_obj_0 <- Surv(time = allnoncarriers$AgeBC, event = allnoncarriers$isAffBC)
  
  km_fit_1 <- survfit(surv_obj_1 ~ 1, data = allcarriers)
  km_fit_0 <- survfit(surv_obj_0 ~ 1, data = allnoncarriers)
  
  # Penetrance curves
  penetrance_1 <- 1 - km_fit_1$surv
  penetrance_0 <- 1 - km_fit_0$surv
  
  # Plotting the curves
  lines(km_fit_1$time, penetrance_1, type = "s", col = colors[1], lty = line_types[1])
  lines(km_fit_0$time, penetrance_0, type = "s", col = colors[2], lty = line_types[2])

  #Add individual points for carriers
  #Align points with step-wise changes by finding the corresponding y-value for each x (age)
  carrier_ages <- allcarriers$AgeBC
  for (age in unique(carrier_ages)) {
    idx <- max(which(km_fit_1$time <= age))
    if(!is.na(idx)) {
      penetrance_at_age <- penetrance_1[idx]
     points(age, penetrance_at_age, col = colors[1], pch = 16)
    }
  }
}

#Plot the empirical penetrance curves
#plot_penetrance_curves(dat)

```

## Estimation using PenEstim

Application of the PenEstim algorithm using the clipp package for the likelihood calculation and the default prior settings.

-   First Quartile \$Q\_{25}\$ & Scaled Beta(6,3)

-   Median \$Q\_{50}\$ & Scaled Beta(2,2)

-   Asymptote \$\gamma\$ & Scaled Beta(1,1)

-   Threshold \$\delta\$ & Uniform(0,25)

Results - Simulated Data - Default Prior

```{r, echo=TRUE, cache=TRUE}

# Load the data for Simulation study 


# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}

# Run Estimation procedure with default prior setting 
out_sim_15k <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 15000, 
    af = 0.9, burn_in = 0.1
)

# print summary stats
out_sim_15k$summary_stats

```

##  New Prior

Here I changed the prior on the asymptote to incorporate an asymptote which is skewed toward later ages.

```{r, echo=TRUE, cache=TRUE}

# Exploring different priors 
prior_params_91 <- list(
    asymptote = list(g1 = 9, g2 = 1),
    shift = list(min = 15, max = 25),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
out_sim_15k_91 <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 15000, 
    prior_params = prior_params_91, af = 0.9, burn_in = 0.1
)

# print summary stats
out_sim_15k_91$summary_stats
plot_traceSingle(out_sim_15k_91$results[[1]])

# Plot the running mean 

runFQ<- running_mean(out_sim_15k_91$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_15k_91$results[[1]]$median_samples)
runShift <- running_mean(out_sim_15k_91$results[[1]]$shift_samples)
runAsy <- running_mean(out_sim_15k_91$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runShift, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Shift Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_15k_91$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_15k_91$results[[1]]$median_samples)
runVShift <- running_variance(out_sim_15k_91$results[[1]]$shift_samples)
runVAsy <- running_variance(out_sim_15k_91$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVShift, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Shift Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")


  # Autocorrelation plots
acf(out_sim_15k_91$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_15k_91$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_15k_91$results[[1]]$shift_samples, main = "ACF of Shift")
acf(out_sim_15k_91$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

```
## Simulation - BRCA1 and Breast Cancer - Additional families

    1.  Updated Prior
    2.  Larger pedigree file with 120 families instead of 60 from the first simulation 
    

```{r, echo=TRUE, cache=TRUE}

# Load the additonal families
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/peelingpearingestim/simFamilies_C_1000_nocen_selected300.RData")

dat<- simFamilies_C_1000_nocen_selected300

# print the description of the families
print(describeFamilies(dat,"Breast","BRCA1"))


# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}


# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 1, g2 = 1),
    shift = list(min = 15, max = 25),
    median = list(m1 = 6, m2 = 3),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
out_sim_10k_simC300 <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 1000, 
    prior_params = prior_params_BRCA1, af = 0.9, burn_in = 0.1, sex = "Female"
)

# print summary stats
out_sim_10k_simC300$summary_stats
plot_traceSingle(out_sim_10k_simC300$results[[1]])

# Plot the running mean 

runFQ<- running_mean(out_sim_10k_simC300$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_10k_simC300$results[[1]]$median_samples)
runShift <- running_mean(out_sim_10k_simC300$results[[1]]$shift_samples)
runAsy <- running_mean(out_sim_10k_simC300$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runShift, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Shift Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_10k_simC300$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_10k_simC300$results[[1]]$median_samples)
runVShift <- running_variance(out_sim_10k_simC300$results[[1]]$shift_samples)
runVAsy <- running_variance(out_sim_10k_simC300$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVShift, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Shift Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")


  # Autocorrelation plots
acf(out_sim_10k_simC300$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_10k_simC300$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_10k_simC300$results[[1]]$shift_samples, main = "ACF of Shift")
acf(out_sim_10k_simC300$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

# plot the scatterplot with the median and first quartile
plot(out_sim_10k_simC300$combined_chains$median_results, out_sim_10k_simC300$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```

## Simulation - Fixed Asymptote

Same code as above but now the asymptote is drawn from a beta with all point mass on 1. 

```{r, echo=TRUE, cache=TRUE}

# Load the additonal families
load("/Users/nicolaskubista/Documents/Master Statistics/Master Thesis/Code/peelingpearingestim/simFamilies_C_1000_nocen_selected300.RData")

dat<- simFamilies_C_1000_nocen_selected300

# print the description of the families
print(describeFamilies(dat,"Breast","BRCA1"))


# Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "PedigreeID"] <- "FamilyID"
    }
}

#  Data Prep
for (i in seq_along(dat)) {
    if ("ID" %in% colnames(dat[[i]])) {
        colnames(dat[[i]])[colnames(dat[[i]]) == "ID"] <- "SubjectID"
    }
}

for (i in seq_along(dat)) {
    # Add a new column "PedigreeID" with the list number
    dat[[i]]$PedigreeID <- i
}


# Exploring different priors 
prior_params_BRCA1 <- list(
    asymptote = list(g1 = 10^10, g2 = 1e-10),
    shift = list(min = 15, max = 25),
    median = list(m1 = 6, m2 = 3),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
# Main Estimation for Female
out_sim_10k_simC300 <- PenEstim(
    data = dat,
    cancer_type = "Breast", gene_input = "BRCA1", n_chains = 4, n_iter_per_chain = 10000, 
    prior_params = prior_params_BRCA1, af = 0.9, burn_in = 0.1, sex = "Female"
)

# print summary stats
out_sim_10k_simC300$summary_stats
plot_traceSingle(out_sim_10k_simC300$results[[1]])

# Plot the running mean 

runFQ<- running_mean(out_sim_10k_simC300$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_10k_simC300$results[[1]]$median_samples)
runShift <- running_mean(out_sim_10k_simC300$results[[1]]$shift_samples)
runAsy <- running_mean(out_sim_10k_simC300$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runShift, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Shift Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_10k_simC300$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_10k_simC300$results[[1]]$median_samples)
runVShift <- running_variance(out_sim_10k_simC300$results[[1]]$shift_samples)
runVAsy <- running_variance(out_sim_10k_simC300$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVShift, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Shift Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")


  # Autocorrelation plots
acf(out_sim_10k_simC300$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_10k_simC300$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_10k_simC300$results[[1]]$shift_samples, main = "ACF of Shift")
acf(out_sim_10k_simC300$results[[1]]$asymptote_samples, main = "ACF of Asymptote")

# plot the scatterplot with the median and first quartile
plot(out_sim_10k_simC300$combined_chains$median_results, out_sim_10k_simC300$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin
```


# Application - CoH Data (Hispanic Cohort)

## MLH6 and Colorectal Cancer

```{r, echo=TRUE, cache=TRUE}

load("/Users/nicolaskubista/Dropbox (Partners HealthCare)/CCGCRN Hispanic Cohort Data/PenEstim/Data/carrierProbandFamilies_cohPedigree_MSH6_COL.RData")

describeFamilies(carrierProbandFamilies_cohPedigree_MSH6_COL,"Colorectal","MSH6")

# Exploring different priors 
prior_params_MSH6 <- list(
    asymptote = list(g1 = 1, g2 = 1),
    shift = list(min = 0, max = 25),
    median = list(m1 = 1, m2 = 1),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
out_sim_15k_MSH6 <- PenEstim(
    data = carrierProbandFamilies_cohPedigree_MSH6_COL,
    cancer_type = "Colorectal", gene_input = "MSH6", n_chains = 4, n_iter_per_chain = 15000,
    prior_params = prior_params_MSH6,
    af = PanelPRODatabase$AlleleFrequency[["MSH6_anyPV",1]], burn_in = 0.1
)

# print summary stats
out_sim_15k_MSH6$summary_stats
plot_traceSingle(out_sim_15k_MSH6$results[[1]])

# Plot the running mean 

runFQ<- running_mean(out_sim_15k_MSH6$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_15k_MSH6$results[[1]]$median_samples)
runShift <- running_mean(out_sim_15k_MSH6$results[[1]]$shift_samples)
runAsy <- running_mean(out_sim_15k_MSH6$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runShift, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Shift Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_15k_MSH6$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_15k_MSH6$results[[1]]$median_samples)
runVShift <- running_variance(out_sim_15k_MSH6$results[[1]]$shift_samples)
runVAsy <- running_variance(out_sim_15k_MSH6$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVShift, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Shift Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")

  # Autocorrelation plots
acf(out_sim_15k_MSH6$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_15k_MSH6$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_15k_MSH6$results[[1]]$shift_samples, main = "ACF of Shift")
acf(out_sim_15k_MSH6$results[[1]]$asymptote_samples, main = "ACF of Asymptote")
```

## MLH1 and Colorectal Cancer

```{r, echo=TRUE, cache=TRUE}

load("/Users/nicolaskubista/Dropbox (Partners HealthCare)/CCGCRN Hispanic Cohort Data/PenEstim/Data/carrierProbandFamilies_cohPedigree_MLH1.RData")

describeFamilies(carrierProbandFamilies_cohPedigree_MLH1,"Colorectal","MLH1")

# Exploring different priors 
prior_params_MLH1 <- list(
    asymptote = list(g1 = 1, g2 = 1),
    shift = list(min = 0, max = 30),
    median = list(m1 = 2, m2 = 2),
    first_quartile = list(q1 = 6, q2 = 3)
)

# Run Estimation procedure with default prior setting 
out_sim_15k_MLH1 <- PenEstim(
    data = carrierProbandFamilies_cohPedigree_MLH1,
    cancer_type = "Colorectal", gene_input = "MLH1", n_chains = 4, n_iter_per_chain = 15000, 
    prior_params = prior_params_MLH1,
    af = PanelPRODatabase$AlleleFrequency[["MLH1_anyPV",1]], burn_in = 0.1
)

# print summary stats
out_sim_15k_MLH1$summary_stats
plot_traceSingle(out_sim_15k_MLH1$results[[1]])

# Plot the running mean 

runFQ<- running_mean(out_sim_15k_MLH1$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_15k_MLH1$results[[1]]$median_samples)
runShift <- running_mean(out_sim_15k_MLH1$results[[1]]$shift_samples)
runAsy <- running_mean(out_sim_15k_MLH1$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runShift, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Shift Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_15k_MLH1$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_15k_MLH1$results[[1]]$median_samples)
runVShift <- running_variance(out_sim_15k_MLH1$results[[1]]$shift_samples)
runVAsy <- running_variance(out_sim_15k_MLH1$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVShift, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Shift Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")

  # Autocorrelation plots
acf(out_sim_15k_MLH1$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_15k_MLH1$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_15k_MLH1$results[[1]]$shift_samples, main = "ACF of Shift")
acf(out_sim_15k_MLH1$results[[1]]$asymptote_samples, main = "ACF of Asymptote")


```


## MSH6 and Colorectal

Here we added are more skewed prior on the median 

```{r, echo=TRUE, cache=TRUE}

# Exploring different priors 
prior_params_MSH6 <- list(
    asymptote = list(g1 = 1, g2 = 1),
    shift = list(min = 0, max = 25),
    median = list(m1 = 8, m2 = 2), # this is new 
    first_quartile = list(q1 = 6, q2 = 3)
)

# New
out_sim_10k_MSH6_newPriorMP <- PenEstim(
    data = carrierProbandFamilies_cohPedigree_MSH6_COL,
    cancer_type = "Colorectal", gene_input = "MSH6", n_chains = 4, n_iter_per_chain = 10000, median_max = FALSE,
    prior_params = prior_params_MSH6, max_penetrance =1,
    af = PanelPRODatabase$AlleleFrequency[["MSH6_anyPV",1]], burn_in = 0.1
)

# print summary stats
out_sim_10k_MSH6_newPriorMP$summary_stats
plot_traceSingle(out_sim_10k_MSH6_newPriorMP$results[[1]])
plot_penetrance(out_sim_10k_MSH6_newPriorMP$combined_chains)
generate_density_plots(out_sim_10k_MSH6_newPriorMP$combined_chains)

# Plot the running mean 

runFQ<- running_mean(out_sim_10k_MSH6_newPriorMP$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_10k_MSH6_newPriorMP$results[[1]]$median_samples)
runShift <- running_mean(out_sim_10k_MSH6_newPriorMP$results[[1]]$shift_samples)
runAsy <- running_mean(out_sim_10k_MSH6_newPriorMP$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runShift, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Shift Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_10k_MSH6_newPriorMP$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_10k_MSH6_newPriorMP$results[[1]]$median_samples)
runVShift <- running_variance(out_sim_10k_MSH6_newPriorMP$results[[1]]$shift_samples)
runVAsy <- running_variance(out_sim_10k_MSH6_newPriorMP$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVShift, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Shift Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")

  # Autocorrelation plots
acf(out_sim_10k_MSH6_newPriorMP$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_10k_MSH6_newPriorMP$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_10k_MSH6_newPriorMP$results[[1]]$shift_samples, main = "ACF of Shift")
acf(out_sim_10k_MSH6_newPriorMP$results[[1]]$asymptote_samples, main = "ACF of Asymptote")


# plot the scatterplot with the median and first quartile
plot(out_sim_10k_MSH6_newPriorMP$combined_chains$median_results, out_sim_10k_MSH6_newPriorMP$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin

```

## MSH6 and Colorectal

    1.  New Prior

    2.  more itterations

```{r, echo=TRUE, cache=TRUE}

# Exploring different priors 
prior_params_MSH6_2 <- list(
    asymptote = list(g1 = 1, g2 = 1),
    shift = list(min = 0, max = 40), # this is new
    median = list(m1 = 6, m2 = 3), # this is new 
    first_quartile = list(q1 = 6, q2 = 3)
)

# New
out_sim_10k_MSH6_newPrior2MP <- PenEstim(
    data = carrierProbandFamilies_cohPedigree_MSH6_COL,
    cancer_type = "Colorectal", gene_input = "MSH6", n_chains = 4, n_iter_per_chain = 10000, median_max = FALSE,
    prior_params = prior_params_MSH6_2, max_penetrance = 0.3,
    af = PanelPRODatabase$AlleleFrequency[["MSH6_anyPV",1]], burn_in = 0.1
)

# print summary stats
out_sim_10k_MSH6_newPrior2MP$summary_stats
plot_traceSingle(out_sim_10k_MSH6_newPrior2MP$results[[1]])
plot_penetrance(out_sim_10k_MSH6_newPrior2MP$combined_chains)
generate_density_plots(out_sim_10k_MSH6_newPrior2MP$combined_chains)

# Plot the running mean 

runFQ<- running_mean(out_sim_10k_MSH6_newPrior2MP$results[[1]]$first_quartile_samples)
runMed <- running_mean(out_sim_10k_MSH6_newPrior2MP$results[[1]]$median_samples)
runShift <- running_mean(out_sim_10k_MSH6_newPrior2MP$results[[1]]$shift_samples)
runAsy <- running_mean(out_sim_10k_MSH6_newPrior2MP$results[[1]]$asymptote_samples)

# Plot running mean
plot(runFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of First Quartile Samples")
plot(runMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Median Samples")
plot(runShift, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Shift Samples")
plot(runAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Mean of Asymptote Samples")


# Plot the running var 

runVFQ<- running_variance(out_sim_10k_MSH6_newPrior2MP$results[[1]]$first_quartile_samples)
runVMed <- running_variance(out_sim_10k_MSH6_newPrior2MP$results[[1]]$median_samples)
runVShift <- running_variance(out_sim_10k_MSH6_newPrior2MP$results[[1]]$shift_samples)
runVAsy <- running_variance(out_sim_10k_MSH6_newPrior2MP$results[[1]]$asymptote_samples)

# Plot running var
plot(runVFQ, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of First Quartile Samples")
plot(runVMed, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Median Samples")
plot(runVShift, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Shift Samples")
plot(runVAsy, type = "l", xlab = "Iteration", ylab = "Running Mean", main = "Running Var of Asymptote Samples")

  # Autocorrelation plots
acf(out_sim_10k_MSH6_newPrior2MP$results[[1]]$first_quartile_samples, main = "ACF of First Quartile")
acf(out_sim_10k_MSH6_newPrior2MP$results[[1]]$median_samples, main = "ACF of Median")
acf(out_sim_10k_MSH6_newPrior2MP$results[[1]]$shift_samples, main = "ACF of Shift")
acf(out_sim_10k_MSH6_newPrior2MP$results[[1]]$asymptote_samples, main = "ACF of Asymptote")


# plot the scatterplot with the median and first quartile
plot(out_sim_10k_MSH6_newPrior2MP$combined_chains$median_results, out_sim_10k_MSH6_newPrior2MP$combined_chains$first_quartile_results, xlim = c(1,100), ylim = c(1,100))
abline(a=0, b=1, col="red") # This draws a red line with slope 1 passing through the origin

```


# Penetrance Estimates from PanelPRO

As a reference, I added the penetrance curves for the respective gene-cancer combinations from the PanelPRODatabase here.

```{r, echo=TRUE, cache=TRUE}
db <- PPP::PanelPRODatabase
ages <- 1:94
# Extracting penetrance data for females
pen1_f <- cumsum(db$Penetrance["Breast", "BRCA1_hetero_anyPV", "All_Races", "Female", ,"Net"])

# Assuming similar structure, extracting for males (modify according to your data structure)
pen1_m <- cumsum(db$Penetrance["Breast", "BRCA1_hetero_anyPV", "All_Races", "Male", ,"Net"])

# Plot for BRCA1 and Breast Cancer including both genders
plot(ages, pen1_f, type="l", lwd=2, col="blue",
     xlab="Age", ylab="Cumulative Penetrance", 
     main="BRCA1 and Breast Cancer")
lines(ages, pen1_m, type="l", lwd=2, col="red")
legend("topleft", legend=c("Female", "Male"), col=c("blue", "red"), lty=1, lwd=2)

# MLH1 and Colorectal Cancer for both genders
pen2_f <- cumsum(db$Penetrance["Colorectal", "MLH1_hetero_anyPV", "All_Races", "Female", ,"Net"])
pen2_m <- cumsum(db$Penetrance["Colorectal", "MLH1_hetero_anyPV", "All_Races", "Male", ,"Net"])

# Plot for MLH1 and Colorectal Cancer
plot(ages, pen2_f, type="l", lwd=2, col="blue",
     xlab="Age", ylab="Cumulative Penetrance", 
     main="MLH1 and Colorectal Cancer")
lines(ages, pen2_m, type="l", lwd=2, col="red")
legend("topleft", legend=c("Female", "Male"), col=c("blue", "red"), lty=1, lwd=2)

# MSH6 and Colorectal Cancer for both genders
pen3_f <- cumsum(db$Penetrance["Colorectal", "MSH6_hetero_anyPV", "All_Races", "Female", ,"Net"])
pen3_m <- cumsum(db$Penetrance["Colorectal", "MSH6_hetero_anyPV", "All_Races", "Male", ,"Net"])

# Plot for MSH6 and Colorectal Cancer
plot(ages, pen3_f, type="l", lwd=2, col="blue",
     xlab="Age", ylab="Cumulative Penetrance", 
     main="MSH6 and Colorectal Cancer")
lines(ages, pen3_m, type="l", lwd=2, col="red")
legend("topleft", legend=c("Female", "Male"), col=c("blue", "red"), lty=1, lwd=2)

```



